---
title: "Sample Size for Mystery Caller Studies"
subtitle: "A simulation-based vignette using Negative Binomial mixed models with overdispersion as the default assumption"
author: "Isochrones Project"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    theme: flatly
    df_print: kable
  pdf_document:
    toc: true
    number_sections: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.align = "center", fig.width = 8, fig.height = 5.5,
  dpi = 150
)
```

# Abstract

This vignette explains how to compute sample sizes for **mystery caller studies**
where each provider is called twice (once with Blue Cross / Blue Shield insurance,
once with Medicaid) and the primary outcome is the **number of business days until
the next available new-patient appointment**. The model is a Negative Binomial
generalized linear mixed model (NB2 GLMM) with a physician-level random intercept
to account for paired calls per provider. Overdispersion is treated as the default
state of the data, not a special case. Two estimands are evaluated: the conditional
rural-by-insurance interaction and the population-marginal Medicaid effect after
post-stratification weighting. A worked example, all model equations, and the full
power simulation are provided.

# Glossary of abbreviations

| Abbreviation | Meaning |
|---|---|
| NPI | National Provider Identifier (10-digit federal physician ID) |
| BCBS | Blue Cross / Blue Shield (commercial insurance signal) |
| GLM | Generalized Linear Model |
| GLMM | Generalized Linear Mixed Model |
| LMM | Linear Mixed Model |
| NB | Negative Binomial distribution |
| NB1 | Negative Binomial with variance $\phi\mu$ (linear in the mean) |
| NB2 | Negative Binomial with variance $\mu + \mu^{2}/\phi$ (quadratic in the mean) |
| OLRE | Observation-Level Random Effect (a per-row random intercept used as a Poisson overdispersion correction) |
| MDE | Minimum Detectable Effect (the smallest true effect the study is powered to find) |
| MLE | Maximum Likelihood Estimation |
| SE | Standard Error |
| CI | Confidence Interval |
| SD | Standard Deviation |
| ICC | Intraclass Correlation Coefficient |
| IRR | Incidence Rate Ratio (exponentiated coefficient in a log-link count model) |
| IPW | Inverse Probability Weight |
| ANOVA | Analysis of Variance |
| RUCA | Rural-Urban Commuting Area code (US Department of Agriculture classification) |
| DGP | Data Generating Process (the simulation truth) |
| MC | Monte Carlo (here, the simulation replicate count) |
| MoE | Margin of Error |

# Study design

* **Sampling units.** $K$ providers identified by NPI.
* **Repeated measure.** Each NPI is called twice in a short window:
  one call presents the patient as **BCBS-insured**; the other as
  **Medicaid-insured**. The two calls share the same provider, so they are
  not independent.
* **Stratification.** Providers are sampled stratified by rurality (rural vs.
  urban based on RUCA code at the practice ZIP). The natural population is
  approximately 15% rural; the analysis below also considers an oversampled
  50/50 design.
* **Outcome.** $Y$ = number of business days from the call to the next available
  appointment for a new patient. This is a count outcome, non-negative,
  and routinely overdispersed in published audit studies (variance well above
  the mean, long right tail, occasional clinic-closed zero or very large values).

The pairing is the central feature of the design. By calling the same NPI twice,
the BCBS-vs-Medicaid contrast is **within-physician**, which removes between-
physician variation from the relevant standard error.

# Statistical model

## Fixed effects (the truth we want to estimate)

Let $i$ index the NPI and $j$ index the call ($j = 1$ for BCBS, $j = 2$ for
Medicaid). Let

- $M_{ij} = 1$ if call $(i,j)$ uses Medicaid, $0$ otherwise.
- $R_{i}  = 1$ if NPI $i$ is rural, $0$ otherwise.

The linear predictor on the log scale is

$$
\eta_{ij} \;=\; \beta_{0} \;+\; \beta_{M}\,M_{ij} \;+\; \beta_{R}\,R_{i} \;+\; \beta_{MR}\,M_{ij} R_{i} \;+\; u_{i}.
$$

where $u_{i} \sim \mathcal{N}(0, \sigma^{2}_{\text{NPI}})$ is the
**physician random intercept**. This intercept absorbs the part of the wait
time that is constant within a provider (busy clinic vs. quiet clinic) and
makes the BCBS vs. Medicaid comparison properly within-NPI.

Exponentiating gives the four cell means in days:

$$
\begin{aligned}
\text{BCBS, urban:}     & \quad \mathbb{E}[Y\mid M=0,R=0] = e^{\beta_{0}} \\
\text{Medicaid, urban:} & \quad e^{\beta_{0}+\beta_{M}} \\
\text{BCBS, rural:}     & \quad e^{\beta_{0}+\beta_{R}} \\
\text{Medicaid, rural:} & \quad e^{\beta_{0}+\beta_{M}+\beta_{R}+\beta_{MR}}
\end{aligned}
$$

Exponentiated coefficients are **incidence rate ratios (IRRs)**:
$e^{\beta_{M}}$ is the Medicaid-vs-BCBS rate ratio for an urban physician.

## Why a Negative Binomial likelihood, not Poisson

A Poisson likelihood enforces $\mathrm{Var}(Y\mid \mu)=\mu$. Real wait-time
counts almost always have variance well above their mean: a few clinics
respond same-week, most respond in a few weeks, and a long right tail of
clinics quote 8+ weeks. The NB likelihood relaxes this constraint. We use
**NB2** (the default in `glmmTMB::nbinom2`):

$$
Y \mid \mu \;\sim\; \mathrm{NB}\!\left(\mu, \phi\right), \qquad
\mathbb{E}[Y\mid\mu]=\mu, \quad
\mathrm{Var}[Y\mid\mu]=\mu + \frac{\mu^{2}}{\phi}.
$$

As $\phi \to \infty$ the NB2 collapses to a Poisson. Smaller $\phi$ means more
overdispersion. With baseline mean wait $\mu_{0}=14$ days and $\phi=6$ the
implied dispersion ratio is $1 + \mu_{0}/\phi = 1 + 14/6 \approx 3.3$, which
matches what we see in mystery caller datasets.

The full GLMM is

$$
\boxed{ \;
Y_{ij} \mid u_{i} \;\sim\; \mathrm{NB2}\!\left(\mu_{ij},\, \phi\right), \qquad
\log \mu_{ij} \;=\; \mathbf{x}_{ij}^{\top}\boldsymbol{\beta} + u_{i}, \qquad
u_{i} \sim \mathcal{N}(0, \sigma^{2}_{\text{NPI}})
\;}
$$

## When to use NB instead of Poisson, and when an OLRE will do

In every analysis, first fit a plain Poisson GLM or GLMM and compute the
**Pearson dispersion ratio**:

$$
\hat{\varphi} \;=\; \frac{1}{n-p}\sum_{ij}\frac{(y_{ij}-\hat{\mu}_{ij})^{2}}{\hat{\mu}_{ij}}.
$$

| $\hat{\varphi}$ | Recommended model |
|---|---|
| 0.8 to 1.2 | Poisson is adequate. |
| 1.2 to 1.5 | Poisson with sandwich (robust) standard errors. |
| 1.5 to 3.0 | NB GLMM or Poisson GLMM with an Observation-Level Random Effect (OLRE). |
| above 3.0 | NB GLMM strongly preferred (OLRE convergence becomes unstable). |

OLRE means adding $(1 \mid \text{obs\_id})$ where `obs_id` is a unique index
per row; the per-row random intercept absorbs extra variance. It is the
standard `lme4` workaround when one cannot use `glmmTMB`. In our setting we
recommend NB directly because (a) `glmmTMB` is fast and stable on this design,
and (b) the variance assumption $\mu + \mu^{2}/\phi$ is more realistic for
right-skewed wait time counts than the OLRE's lognormal mixing.

# Estimands and post-stratification weighting

We care about two different effects from the same model.

## Conditional rural-by-insurance interaction

The coefficient $\beta_{MR}$ in the equation above. On the log scale it is the
ratio of two rate ratios:

$$
\beta_{MR} \;=\; \log \frac{\mathrm{IRR}_{\text{Medicaid vs BCBS, rural}}}
                            {\mathrm{IRR}_{\text{Medicaid vs BCBS, urban}}}.
$$

A Wald $z$ test on $\beta_{MR}$ answers: *does the Medicaid penalty differ
between rural and urban providers?* This test does **not** need any weighting
because it conditions on rurality.

## Population-marginal Medicaid effect

We also want a single, generalizable number for the Medicaid penalty across
the whole physician population, on the natural (days) scale. Define

$$
\Delta_{\text{pop}}
\;=\;
p_{R}\cdot\!\left(\mathbb{E}[Y\mid M=1,R=1]-\mathbb{E}[Y\mid M=0,R=1]\right)
\;+\;
(1-p_{R})\cdot\!\left(\mathbb{E}[Y\mid M=1,R=0]-\mathbb{E}[Y\mid M=0,R=0]\right)
$$

where $p_{R}$ is the population rural fraction (we use $p_{R}=0.15$).

If we stratify-sample 50/50 rural-urban for power efficiency, the **raw**
average Medicaid effect from our sample is biased upward, because rural
contributes 50% of the average instead of 15%. We correct this with **inverse-
probability post-stratification weights**:

$$
w_{ij} \;=\;
\begin{cases}
\dfrac{p_{R}}{q_{R}}             & R_{i}=1 \quad \text{(rural)} \\[6pt]
\dfrac{1-p_{R}}{1-q_{R}}         & R_{i}=0 \quad \text{(urban)}
\end{cases}
$$

where $q_{R}$ is the **sampling** rural fraction. For a 50/50 design with
$p_{R}=0.15$: rural rows get $w=0.30$ and urban rows get $w=1.70$, so they
reproduce the 15/85 population mix.

The weighted marginal effect from `marginaleffects::avg_comparisons()` then
targets $\Delta_{\text{pop}}$ unbiasedly. With our cell means
(BCBS-urban = 14, Med-urban = 17, BCBS-rural = 18, Med-rural = 24):

$$
\Delta_{\text{pop}}
= 0.15\cdot(24-18) + 0.85\cdot(17-14)
= 0.90 + 2.55
= 3.45 \;\text{days}.
$$

# Simulation methodology

Power is the probability that the study correctly rejects the null hypothesis
of no effect. With a GLMM we approximate it by Monte Carlo:

1. Set the truth: cell means, $\sigma_{\text{NPI}}$, $\phi$, $p_{R}$, $q_{R}$.
2. Build a design grid of $K$ NPIs with random rural assignment at rate $q_{R}$
   and two rows per NPI (BCBS, Medicaid).
3. For each Monte Carlo replicate $m = 1,\ldots,M$:
   - Draw NPI random intercepts $u_{i}^{(m)} \sim \mathcal{N}(0,\sigma_{\text{NPI}}^{2})$.
   - Compute $\mu_{ij}^{(m)} = \exp(\mathbf{x}_{ij}^{\top}\boldsymbol{\beta} + u_{i}^{(m)})$.
   - Draw outcomes $Y_{ij}^{(m)} \sim \mathrm{NB2}(\mu_{ij}^{(m)}, \phi)$.
   - Fit `glmmTMB(Y ~ insurance * rural + (1 | npi), family = nbinom2)`.
   - Record (i) the Wald $p$-value for $\beta_{MR}$ and (ii) the post-stratified
     marginal Medicaid effect via `marginaleffects::avg_comparisons()` with
     `wts = "popwt"`.
4. Estimated power at sample size $K$ is the proportion of replicates with
   $p < \alpha$ for each estimand.

The full driver script is in `scripts/mystery_caller_power_NB.R`.

# Worked example: the realistic recruitment target

We ran the simulation under these settings:

- BCBS-urban = 14, Medicaid-urban = 17, BCBS-rural = 18, Medicaid-rural = 24 days
- $\sigma_{\text{NPI}} = 0.30$, $\phi = 6$ (overdispersion ratio about 3.3 at $\mu = 14$)
- $\alpha = 0.05$, two-sided
- 50/50 stratified rural-urban sampling, population $p_{R} = 0.15$
- $M = 30$ Monte Carlo replicates per sample size

```{r results-table, echo = FALSE}
res <- utils::read.csv(file.path("..", "artifacts", "power_analysis",
                                 "nb_power_50_50.csv"))
res$pow_cond_interact <- sprintf("%.0f%%", 100 * res$pow_cond_interact)
res$pow_marg_unwtd    <- sprintf("%.0f%%", 100 * res$pow_marg_unwtd)
res$pow_marg_popwtd   <- sprintf("%.0f%%", 100 * res$pow_marg_popwtd)
res$mean_marg_pop_est <- sprintf("%.2f", res$mean_marg_pop_est)
colnames(res) <- c("NPIs", "Calls",
                   "Power: cond. interaction",
                   "Power: marg., unweighted",
                   "Power: marg., pop-weighted",
                   "Est. marginal effect (days)",
                   "Monte Carlo replicates")
knitr::kable(res, align = "r",
             caption = "Power by sample size, NB2 GLMM, 50/50 stratified design.")
```

```{r power-curve, echo = FALSE, fig.cap = "Power curves for the three estimands as a function of total phone calls. Dashed line at 90 percent. The two marginal estimands hit ceiling power almost immediately; the conditional interaction is the binding constraint."}
knitr::include_graphics(
  file.path("..", "artifacts", "power_analysis", "figures",
            "mystery_caller_power_curves.png")
)
```

## Interpretation

- **Marginal Medicaid effect** (both unweighted and population-weighted) is
  detected at 100% power even at $K=200$. The marginal contrast is a
  big effect (~3.45 days) with a small standard error, so it is easy to
  detect. The weighted version recovers the true 3.45-day value; the
  unweighted version is biased high (about 4.5 days) because the 50/50
  sample over-represents rural.

- **Rural-by-insurance interaction** is the binding constraint. It requires
  about $K = 2{,}000$ NPIs (4,000 calls) to reach 90% power under NB2
  with the assumed overdispersion. Naive Poisson would suggest about
  $K = 1{,}000$ but its standard errors are too small under overdispersion,
  so its rejections are inflated. The NB answer is the honest one.

## What if you cannot afford 4,000 calls?

Two levers reduce the requirement:

1. **Stratify harder.** Push beyond 50/50 to oversample rural (say 70/30).
   The interaction's standard error is minimized when the rural-urban split
   matches the proportion of the interaction's signal. With more rural in
   the sample, $K$ for 90% power drops.
2. **Drop the interaction question.** If the design only needs to detect a
   marginal Medicaid penalty (no rural-urban differential), the population-
   weighted estimand reaches 90% power at $K \approx 50$ NPIs (about 100
   calls) and 100% at $K \approx 200$.

Concretely:

| Research question | Approximate $K$ for 90% power |
|---|---:|
| Is there a Medicaid penalty at all (marginal)? | ~50 NPIs / ~100 calls |
| Is the Medicaid penalty larger in rural areas (interaction)? | ~2,000 NPIs / ~4,000 calls |

# Why we moved away from `simr`

The `simr` package is excellent for the simple case: one `merMod` object,
one fixed-effect coefficient to test, sweep one variable. It became limiting
once the design grew:

| Feature | `simr` | Custom simulation loop |
|---|---|---|
| Negative Binomial GLMM | Not natively supported (only `lme4` families) | Yes, via `glmmTMB::nbinom2` |
| Test a derived (marginal) estimand | Awkward | Direct, via `marginaleffects` |
| Apply inverse-probability weights | Awkward | Trivial, `wts = "popwt"` |
| Test several estimands per replicate | One at a time | All in one pass |
| Speed | Wraps `lme4` + Kenward-Roger | Direct, easy to parallelize |
| Custom data-generating processes | Limited | Anything you can simulate |

Our recommendation is to keep the simulation loop entirely in our own R code:
`simulate -> fit -> extract estimand -> tally rejections`. The driver is in
`scripts/mystery_caller_power_NB.R` and is ~150 lines.

# A planned Shiny app

For routine use we will wrap this simulation in a Shiny application so that
investigators can pick scenarios interactively. Planned inputs:

- Baseline mean wait (days)
- Hypothesized cell means or a single MDE in days
- NPI random intercept SD (default 0.30 on log scale; if pilot data exist,
  estimate from `glmer(y ~ 1 + (1 | npi))` and use the fitted SD)
- NB dispersion $\phi$ (default 6 for moderate overdispersion)
- Sampling rural fraction $q_{R}$
- Population rural fraction $p_{R}$
- $\alpha$ and target power
- Monte Carlo replicate count (default 100; warning for >300)

Planned outputs:

- Recommended $K$ for each estimand to hit the target power
- Power curve plot
- Sensitivity panel: how $K$ moves as $\phi$, $\sigma_{\text{NPI}}$, and
  $q_{R}$ change
- Downloadable CSV of the simulation grid for inclusion in a protocol

The Shiny app skeleton will live at `inst/shiny/mystery_caller_power/`
once built. Because each Monte Carlo replicate fits a `glmmTMB` model, the
backend will dispatch sims in parallel with `future.apply` and cache
identical scenarios.

# Reproducibility

```{r session-info, echo = FALSE}
sessionInfo()
```

# References

- Brooks ME et al. (2017). glmmTMB Balances Speed and Flexibility Among
  Packages for Zero-inflated Generalized Linear Mixed Models. *R Journal*.
- Bates D, Maechler M, Bolker B, Walker S (2015). Fitting Linear Mixed-
  Effects Models Using lme4. *J. Statistical Software*.
- Arel-Bundock V (2024). `marginaleffects`: Predictions, Comparisons, Slopes,
  Marginal Means, and Hypothesis Tests.
- Lumley T (2010). *Complex Surveys: A Guide to Analysis Using R*.
- Harrison XA (2014). Using observation-level random effects to model
  overdispersion in count data. *PeerJ*.
