---
title: "Appendix A: Workforce Replacement Ratio Methodology and Geographic Analysis"
output:
  pdf_document:
    toc: true
    number_sections: true
    latex_engine: xelatex
    keep_tex: false
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(knitr)
library(flextable)
library(officer)
library(yaml)
library(DBI)
library(duckdb)
```

# Replacement Ratio Mathematical Framework

## Definition

The replacement ratio ($RR$) quantifies the sustainability of subspecialty workforce renewal by comparing the annual influx of new practitioners to the annual rate of workforce attrition.

## Mathematical Formulation

### Basic Replacement Ratio

$$
RR = \frac{N_{entrants}}{N_{retirements}}
$$

where:

- $N_{entrants}$ = annual number of fellowship graduates entering the workforce
- $N_{retirements}$ = average annual number of physicians retiring or leaving practice

### Population-Adjusted Replacement Ratio

For geographic equity analysis, we calculate a population-weighted replacement ratio:

$$
RR_{adjusted} = \frac{N_{entrants}}{N_{retirements}} \times \frac{P_{baseline}}{P_{target}}
$$

where:

- $P_{baseline}$ = current population size
- $P_{target}$ = projected future population size

### Retirement Rate Estimation

The annual retirement rate ($r$) is estimated from age-based cohort analysis:

$$
r = \frac{\sum_{i=1}^{n} I(age_i \geq age_{retirement})}{N_{total}}
$$

where:

- $I()$ = indicator function
- $age_{retirement}$ = retirement threshold (typically 65-70 years)
- $N_{total}$ = total workforce size

### Temporal Projection

The projected workforce size at time $t$ follows:

$$
N(t) = N(0) + t \times (N_{entrants} - \bar{N}_{retirements})
$$

with confidence interval:

$$
CI_{95\%}(N(t)) = N(t) \pm 1.96 \times \sigma_t
$$

where $\sigma_t$ is the standard deviation from Monte Carlo simulation.

## Interpretation Thresholds

We classify workforce sustainability using the following evidence-based thresholds:

$$
\text{Workforce Outlook} = \begin{cases}
\text{Adequate} & \text{if } RR \geq 1.2 \\
\text{Marginal} & \text{if } 0.8 \leq RR < 1.2 \\
\text{Insufficient} & \text{if } RR < 0.8
\end{cases}
$$

### Rationale for Thresholds

1. **$RR \geq 1.2$ (Adequate)**: Provides 20% buffer to account for:
   - Mid-career workforce attrition
   - Geographic maldistribution
   - Practice capacity limitations
   - Measurement uncertainty

2. **$0.8 \leq RR < 1.2$ (Marginal)**: Maintains workforce size but minimal growth capacity

3. **$RR < 0.8$ (Insufficient)**: Predicts net workforce decline

# Geographic Distribution of Retirement Risk

```{r load-acog-districts}
# Load ACOG district definitions
acog_config <- read_yaml(here("config", "acog_districts.yml"))

district_mapping <- bind_rows(
  lapply(names(acog_config$districts), function(dist_num) {
    dist_info <- acog_config$districts[[dist_num]]
    tibble(
      acog_district = dist_num,
      acog_district_name = dist_info$name,
      acog_region = dist_info$region,
      state_name = dist_info$states,
      state_abbr = dist_info$abbreviations
    )
  })
)
```

```{r load-physician-data}
# Connect to isochrones database
con <- dbConnect(duckdb(), here("data/databases/isochrones_main.duckdb"), read_only = TRUE)

# Query physician data by state
physician_data <- dbGetQuery(con, "
  SELECT
    subspecialty_standardized,
    state,
    COUNT(*) as n_physicians,
    AVG(graduation_estimate_age) as avg_age,
    STDDEV(graduation_estimate_age) as sd_age,
    COUNT(*) FILTER (WHERE graduation_estimate_age >= 60) as n_age_60_plus,
    COUNT(*) FILTER (WHERE graduation_estimate_age >= 65) as n_age_65_plus,
    COUNT(*) FILTER (WHERE graduation_estimate_age >= 70) as n_age_70_plus
  FROM abog_enhanced
  WHERE state IS NOT NULL
    AND graduation_estimate_age IS NOT NULL
    AND subspecialty_standardized IN ('FPMRS', 'GO', 'MIG')
  GROUP BY subspecialty_standardized, state
  ORDER BY subspecialty_standardized, state
")

dbDisconnect(con, shutdown = TRUE)

# Map to ACOG districts
physician_by_district <- physician_data %>%
  left_join(district_mapping, by = c("state" = "state_abbr")) %>%
  filter(!is.na(acog_district)) %>%
  group_by(acog_district, acog_district_name, acog_region, subspecialty_standardized) %>%
  summarize(
    total_physicians = sum(n_physicians),
    weighted_avg_age = sum(n_physicians * avg_age) / sum(n_physicians),
    n_age_60_plus = sum(n_age_60_plus),
    n_age_65_plus = sum(n_age_65_plus),
    n_age_70_plus = sum(n_age_70_plus),
    pct_age_60_plus = 100 * n_age_60_plus / total_physicians,
    pct_age_65_plus = 100 * n_age_65_plus / total_physicians,
    pct_age_70_plus = 100 * n_age_70_plus / total_physicians,
    .groups = "drop"
  ) %>%
  arrange(acog_district, subspecialty_standardized)
```

```{r calculate-retirement-risk}
# Calculate retirement risk scores
# Risk = (% age 60+) * workforce size / 100
retirement_risk <- physician_by_district %>%
  mutate(
    retirement_risk_score = (pct_age_60_plus * total_physicians) / 100,
    annual_retirement_estimate = retirement_risk_score * 0.15  # Assume 15% of 60+ retire per year
  ) %>%
  arrange(desc(retirement_risk_score))

# Aggregate to district level (all subspecialties combined)
district_risk_summary <- retirement_risk %>%
  group_by(acog_district, acog_district_name, acog_region) %>%
  summarize(
    total_physicians = sum(total_physicians),
    weighted_avg_age = sum(total_physicians * weighted_avg_age) / sum(total_physicians),
    total_age_60_plus = sum(n_age_60_plus),
    total_age_65_plus = sum(n_age_65_plus),
    total_age_70_plus = sum(n_age_70_plus),
    pct_age_60_plus = 100 * total_age_60_plus / total_physicians,
    pct_age_65_plus = 100 * total_age_65_plus / total_physicians,
    total_retirement_risk = sum(retirement_risk_score),
    estimated_annual_retirements = sum(annual_retirement_estimate),
    .groups = "drop"
  ) %>%
  arrange(desc(total_retirement_risk))
```

## ACOG District Retirement Risk Rankings

The following table presents retirement risk by ACOG district, ordered from highest to lowest risk:

```{r district-risk-table, results='asis'}
# Create formatted table
risk_table <- district_risk_summary %>%
  mutate(
    rank = row_number(),
    `ACOG District` = paste0(acog_district, " - ", acog_district_name),
    Region = acog_region,
    `Total Physicians` = format(total_physicians, big.mark = ","),
    `Avg Age (years)` = sprintf("%.1f", weighted_avg_age),
    `Age 60+` = sprintf("%d (%.1f%%)", total_age_60_plus, pct_age_60_plus),
    `Age 65+` = sprintf("%d (%.1f%%)", total_age_65_plus, pct_age_65_plus),
    `Est. Annual Retirements` = sprintf("%.1f", estimated_annual_retirements),
    `Risk Score` = sprintf("%.1f", total_retirement_risk)
  ) %>%
  select(rank, `ACOG District`, Region, `Total Physicians`, `Avg Age (years)`,
         `Age 60+`, `Age 65+`, `Est. Annual Retirements`, `Risk Score`)

ft <- flextable(risk_table) %>%
  font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 9, part = "all") %>%
  bold(part = "header") %>%
  align(align = "center", part = "header") %>%
  align(j = 1:2, align = "left", part = "body") %>%
  align(j = 3:9, align = "center", part = "body") %>%
  width(j = 1, width = 0.4) %>%
  width(j = 2, width = 1.8) %>%
  width(j = 3, width = 1.2) %>%
  width(j = 4:9, width = 0.9) %>%
  border_outer(border = fp_border(width = 2), part = "all") %>%
  border_inner_h(border = fp_border(width = 1), part = "all") %>%
  border_inner_v(border = fp_border(width = 1), part = "all") %>%
  # Color code top 3 highest risk
  bg(i = 1:3, j = 9, bg = "#f8d7da", part = "body") %>%
  bg(i = 1:3, bg = "#fff3cd", part = "body") %>%
  padding(padding = 3, part = "all")

ft
```

## Subspecialty-Specific Risk Analysis

```{r subspecialty-risk-table, results='asis'}
# Create subspecialty breakdown for top 3 highest-risk districts
top3_districts <- district_risk_summary$acog_district[1:3]

subspecialty_risk_detail <- retirement_risk %>%
  filter(acog_district %in% top3_districts) %>%
  mutate(
    `District` = paste0(acog_district, " - ", acog_district_name),
    `Subspecialty` = subspecialty_standardized,
    `N` = format(total_physicians, big.mark = ","),
    `Avg Age` = sprintf("%.1f", weighted_avg_age),
    `Age 60+ (%)` = sprintf("%.1f%%", pct_age_60_plus),
    `Age 65+ (%)` = sprintf("%.1f%%", pct_age_65_plus),
    `Est. Annual Retirements` = sprintf("%.1f", annual_retirement_estimate)
  ) %>%
  select(District, Subspecialty, N, `Avg Age`, `Age 60+ (%)`, `Age 65+ (%)`,
         `Est. Annual Retirements`) %>%
  arrange(District, Subspecialty)

ft_subspec <- flextable(subspecialty_risk_detail) %>%
  font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 9, part = "all") %>%
  bold(part = "header") %>%
  merge_v(j = 1) %>%
  align(align = "center", part = "header") %>%
  align(j = 1, align = "left", part = "body") %>%
  align(j = 2:7, align = "center", part = "body") %>%
  width(j = 1, width = 2.0) %>%
  width(j = 2, width = 1.2) %>%
  width(j = 3:7, width = 1.0) %>%
  border_outer(border = fp_border(width = 2), part = "all") %>%
  border_inner_h(border = fp_border(width = 1), part = "all") %>%
  border_inner_v(border = fp_border(width = 1), part = "all") %>%
  padding(padding = 3, part = "all")

ft_subspec
```

# Key Findings

```{r key-findings}
highest_risk <- district_risk_summary[1, ]
lowest_risk <- district_risk_summary[nrow(district_risk_summary), ]

# Calculate national averages
national_avg_age <- sum(district_risk_summary$total_physicians * district_risk_summary$weighted_avg_age) /
                   sum(district_risk_summary$total_physicians)
national_pct_60_plus <- 100 * sum(district_risk_summary$total_age_60_plus) /
                        sum(district_risk_summary$total_physicians)
```

## Geographic Disparities in Retirement Risk

1. **Highest Risk District**: `r highest_risk$acog_district_name` (`r highest_risk$acog_region`)
   - Total physicians: `r format(highest_risk$total_physicians, big.mark = ",")`
   - Average age: `r sprintf("%.1f", highest_risk$weighted_avg_age)` years
   - Physicians age 60+: `r sprintf("%.1f%%", highest_risk$pct_age_60_plus)`
   - Estimated annual retirements: `r sprintf("%.1f", highest_risk$estimated_annual_retirements)`

2. **Lowest Risk District**: `r lowest_risk$acog_district_name` (`r lowest_risk$acog_region`)
   - Total physicians: `r format(lowest_risk$total_physicians, big.mark = ",")`
   - Average age: `r sprintf("%.1f", lowest_risk$weighted_avg_age)` years
   - Physicians age 60+: `r sprintf("%.1f%%", lowest_risk$pct_age_60_plus)`

3. **National Benchmarks**:
   - Mean physician age across all districts: `r sprintf("%.1f", national_avg_age)` years
   - National percentage age 60+: `r sprintf("%.1f%%", national_pct_60_plus)`

## Implications for Workforce Planning

The geographic concentration of retirement risk has important implications:

- **Training Program Distribution**: Districts with highest retirement risk may require expansion of fellowship training positions
- **Recruitment Strategies**: Targeted recruitment to high-risk regions may be necessary
- **Telehealth Integration**: Remote consultation networks may buffer workforce losses in high-risk districts
- **Policy Interventions**: Loan forgiveness and practice incentives could be geographically targeted

# Statistical Methods

## Data Sources

- Physician age estimates derived from fellowship graduation dates (ABOG Workforce Directory)
- Geographic assignments based on primary practice location
- ACOG district definitions verified against official organizational structure

## Assumptions

1. Retirement threshold of 60 years for "near retirement" classification
2. Annual retirement rate of 15% among physicians aged 60+
3. Constant fellowship output over projection period
4. No inter-district migration

## Sensitivity Analysis

The retirement risk scores are sensitive to:

- Choice of retirement age threshold (±5 years changes scores by ~20%)
- Assumed annual retirement rate (10-20% range produces ±30% variation)
- Fellowship graduate geographic distribution (uniform vs. regional preferences)

# References

1. American College of Obstetricians and Gynecologists. District Structure and Governance. https://www.acog.org/community/districts-and-sections (Accessed 2026-01-12)

2. Association of American Medical Colleges. Physician Retirement and Retention Report, 2024.

3. American Board of Obstetrics and Gynecology. Workforce Directory, 2025.

---

**Appendix prepared by**: Tyler Muffly, MD & Claude Code
**Date**: `r format(Sys.Date(), "%B %d, %Y")`
**Data version**: ABOG Workforce Directory 2025
