---
title: "Supplemental Digital Content"
subtitle: "Supplemental Digital Content for Projected Clinical-Practice Departure and Fellowship Pipeline Balance in Gynecologic Oncology and Urogynecology, 2025-2029, Including Exploratory MIGS Analyses"
output:
  word_document:
    reference_docx: templates/green_journal_reference.docx
    fig_caption: true
bibliography: references_WORKFORCE_CLIFF.bib
csl: csl/american-medical-association.csl
link-citations: true
---

```{r setup, include=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(here)
  library(knitr)
})
source(here("manuscript/R/workforce_statistics.R"))
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
workforce_summary <- load_workforce_data()
```

# Appendix S1. Supplemental Methods

## Six-source hierarchical clinical-practice departure detection

Clinical-practice departure was classified by integrating six administrative
evidence streams, each assigned a fixed priority (for conflict resolution) and
confidence weight (Appendix Table S2). Higher-priority sources overrode
lower-priority sources when signals disagreed. The confidence weights determine the
consensus departure *year* and a confidence score; they do not set a departure-versus-active
threshold. A candidate departure is detected when at least one qualifying cessation
signal fires (each only for physicians with prior positive evidence in that source) and
no active counter-signal overrides (see the explicit Step 1/Step 2 rule in Appendix S4).
As in the main text, "departure" denotes sustained departure from active clinical
practice, not age-related retirement.

A single uniform three-year cessation rule was applied to both cohorts; no
subspecialty-specific inactivity threshold was used. The full algorithm, including the
exact Step 1/Step 2 Boolean rule, tie-breaking, prior-evidence requirements, and
missing-data handling, is specified in Appendix S4.

## Workforce projection

The active workforce was projected with a dynamic age-structured stock-flow model
over the four-year horizon (2025-2029). The workforce was represented as its
distribution of active physicians by age. In each year every physician aged one
year, the empirical age-band departure hazard was applied to the current age
distribution to yield that year's departures, and that year's fellowship
graduates were added at their entry age; the survivors and new entrants formed
the next year's distribution. Because the cohort ages into higher-hazard bands,
this produces modestly more departures than a static single-rate projection; for
transparency the static projection (2025 baseline plus the four-year net of
graduates and baseline-rate departures) is retained as a benchmark
(Appendix Table S1).

The age-band departure hazards were derived empirically from the study cohort's own
departures over the fully observable window 2016-2021: a person-year life table
produced age-band annual hazards (events divided by person-years at risk within
each band, with age reconstructed from the board-certification year). Because
departure events are sparse within any single subspecialty age band, a single
**pooled** age-band hazard was estimated across the gynecologic surgical
subspecialties and applied to each subspecialty's own active-age distribution; the
per-subspecialty rates differ because of age structure, not separate estimation.
Appendix S6b reports the by-band event and person-year counts, the unpooled
subspecialty-specific hazards, and the pooled hazard, and shows that the pooling
assumption (not separate fitting) determines which subspecialty's rate is higher.
Because the classifier requires a sustained multi-year absence and the data end in
2023-2024, departures dated 2022-2023 are unconfirmable and were treated as
right-censored (they otherwise inflate the rate as provisional over-flags). Applied
to the 2025 active-age distribution the pooled hazard implies departure rates of
0.8% to 1.2% per year, at or below the 1.5% to 1.7% national all-surgeon attrition
benchmark, consistent with these being low-attrition specialties. The URPS baseline
is the both-pathway (ABOG plus ABU) active workforce.

The 2025 active baseline is an estimate carried forward from the most recent
administrative signals (Medicare through 2023, Open Payments through 2024), not a
directly observed 2025 census; it could therefore misclassify physicians who departed
in 2023-2024 (whose sustained absence is not yet confirmable) or omit physicians
entering after the data freeze. To bound this, net annual additions (entrants minus
departures) are about +62 for Gynecologic Oncology and +50 for URPS, so a one- to
two-year baseline lag shifts the 2025 baseline and the 2029 count by at most roughly
60 to 120 (Gynecologic Oncology) and 50 to 100 (URPS), about 5% to 9% of the workforce.
Because recent net flow is positive, an under-counted baseline would make the projected
growth if anything conservative; a retained-departure bias would act in the opposite
direction and partly offset it. The completion-to-departure ratio, a flow balance of entrants to
departures, is unaffected by baseline level in either direction. This baseline-lag
uncertainty is not captured in the parameter-uncertainty interval.

Annual potential entrants were the fellowship completer count, averaged over the
four most recent academic years (2020-2021 to 2023-2024): ACGME Data Resource Book
graduating residents for Gynecologic Oncology (mean 75) and URPS (both-pathway mean
64, obstetrics-gynecology plus urology; Table D.5), and the NRMP/AAGL match fill for
MIGS (47), which is not ACGME-accredited and has no ACGME graduate series. NRMP
positions filled is retained as a benchmark inflow. Because NRMP counts fellows
beginning training, the optimistic scenario lags each 2026 appointment-year matched
cohort by its fellowship duration and adds it to graduate output only in the expected
completion year: Gynecologic Oncology and obstetrics-gynecology-pathway URPS at 3
years (completing 2029), urology-pathway URPS at 2 years (completing 2028); no
separate training-attrition discount is applied to the benchmark, and fellows
completing after 2029 fall outside the horizon. Within the 2025-2029 window the
benchmark therefore raises URPS output in 2028 (urology component) and 2029 (both),
and Gynecologic Oncology output in 2029 only.

Uncertainty in the 2029 count was characterized by a 10,000-iteration Monte Carlo
parameter-uncertainty simulation (seed 20260718) that drew each age-band hazard from its Beta(events+0.5,
person-years-events+0.5) posterior and each subspecialty's annual graduate output
by empirical resampling of the four observed annual completer counts; the 2.5th and 97.5th percentiles are reported. Structural sensitivity
was assessed by varying the observation window (2016-2021, 2016-2019, 2016-2023),
the departure rate (up to three times the empirical estimate), and the fraction of
graduates entering practice (70% to 100%), each passed through the same dynamic model.

## Replacement balance

The completion-to-departure ratio is annual fellowship graduates divided by mean annual
departures, describing headcount replacement balance only. Each cohort was labeled
above replacement (ratio >1.05), at replacement (0.95-1.05), or below replacement
(<0.95); the 20% buffer (`r get_adequate_threshold()`) is retained as a stricter
sensitivity threshold, not an adequacy standard.

# Appendix Table S1. Complete Workforce-Projection Data

```{r table-s1}
s1 <- workforce_summary %>%
  transmute(
    Subspecialty = subspecialty,
    Abbrev = subspecialty_abbrev,
    `Baseline 2025 (n)` = format(baseline_2025, big.mark = ",", trim = TRUE),
    `Projected 2029 (n)` = format(round(projected_2029), big.mark = ",", trim = TRUE),
    `95% interval` = paste0(format(round(ci95_lower), big.mark = ",", trim = TRUE), "-",
                      format(round(ci95_upper), big.mark = ",", trim = TRUE)),
    `Change (%)` = sprintf("%+.1f", percent_change),
    `2025 rate (%)` = sprintf("%.1f", annual_retirement_rate),
    `Mean departures/yr` = sprintf("%.1f", avg_annual_retirements),
    `Graduates/yr` = annual_entrants,
    `Graduates 4yr` = format(annual_entrants * WORKFORCE_PROJECTION_HORIZON_YEARS,
                           big.mark = ",", trim = TRUE),
    `Departures 4yr` = format(round(avg_annual_retirements * WORKFORCE_PROJECTION_HORIZON_YEARS),
                          big.mark = ",", trim = TRUE),
    `Ratio` = sprintf("%.2f", replacement_ratio),
    Assessment = replacement_assessment
  )
kable(s1, caption = "**Appendix Table S1.** Complete workforce-projection data for all three gynecologic surgical subspecialties (2025-2029), dynamic age-structured model.")
```

*Primary workforce-projection values derive from the contract-validated consolidated projection dataset
(`cliff/data/workforce_projections_consolidated.csv`); secondary sensitivity analyses derive from their respective version-controlled analytic outputs. Graduate and departure
totals both span the same four-year projection horizon. Graduates are ACGME
completers (GO, URPS) or the NRMP/AAGL match fill (MIGS, not ACGME-accredited).
The 95% interval is the 2.5th to 97.5th percentile of a 10,000-iteration Monte Carlo
parameter-uncertainty simulation propagating age-band hazard uncertainty (Beta posteriors) and year-to-year
graduate variability (a parameter-uncertainty interval, not a statistical confidence or full predictive interval). URPS, Urogynecology and
Reconstructive Pelvic Surgery; GO, Gynecologic Oncology; MIGS, Minimally Invasive
Gynecologic Surgery.*

# Appendix Table S2. Departure-Detection Sources

```{r table-s2}
s2 <- tibble::tibble(
  Source = c("ABMS Maintenance of Certification (MOC)", "NPPES deactivation",
             "PECOS enrollment", "Open Payments", "Medicare Part D",
             "Medicare Part B"),
  Priority = c(1, 2, 4, 3, 5, 6),
  `Confidence weight` = sprintf("%.2f", c(0.95, 0.90, 0.80, 0.70, 0.60, 0.60)),
  `Departure signal` = c(
    "MOC lapse", "Permanent NPI deactivation without reactivation",
    "Medicare enrollment disenrollment", "Industry payment cessation ≥2 years",
    "Prescribing cessation ≥3 years", "Claims cessation ≥3 years")
) %>% dplyr::arrange(Priority)
kable(s2, caption = "**Appendix Table S2.** Six-source hierarchical clinical-practice departure-detection framework, priority order, and confidence weights.")
```

*ABMS, American Board of Medical Specialties; NPPES, National Plan and Provider
Enumeration System; PECOS, Provider Enrollment, Chain, and Ownership System.*

# Appendix S2. General Obstetrics and Gynecology Comparator and Benchmark

To benchmark the departure signal in a large, well-defined population and to
probe the young MIGS departure ages, we applied the departure-*detection* step (the
Open-Payments-inclusive Step 1, before the non-Open-Payments anchor and before
age-stratification) to a comparator cohort of general obstetrician-gynecologists
(Medicare Doctors and Clinicians file primary specialty "Obstetrics/Gynecology",
excluding every ABOG-certified subspecialist so the comparator is generalist-only).
This is a crude cross-cohort benchmark, NOT the primary classifier: it uses the full
certified cohort (Gynecologic Oncology n = 1,190; URPS n = 1,135) rather than the
anchored active baseline, a crude non-age-stratified rate rather than the pooled
age-band hazard, Open-Payments-inclusive detection rather than the non-Open-Payments
anchor, and Healthgrades profile age rather than the certification-anchored age; its
departure counts and median ages therefore differ from the primary Table 3 (37 and 35
anchored departures; median age 58 and 47) and are not directly comparable to them.
Subspecialty was defined by ABOG board certification, not by the administrative
specialty code, which lumps all OB/GYN subspecialists under a single generic value.
This is a benchmark and negative control, not an individual-level validation, and the
generalists are not pooled with the subspecialty projections; the pooled age-band
hazard applied to specialty-specific age distributions remains the primary model input.

Age at departure used the physicians' actual Healthgrades profile age for the
ABOG subspecialists (Healthgrades does not scrape a generalist roster) and the
graduation-year proxy for the general cohort; a single-source graduation-year
column is shown for a like-for-like comparison across all four cohorts.

```{r table-s3}
s3 <- readr::read_csv(here::here("cliff","data","general_obgyn_comparator.csv"),
                      show_col_types = FALSE)
s3v <- s3 %>%
  transmute(
    Cohort = cohort,
    `N` = format(n, big.mark = ",", trim = TRUE),
    `Annual departure rate (%)` = sprintf("%.2f", annual_departure_rate_pct),
    `Age source` = age_source,
    `Median departure age` = median_departure_age,
    `IQR` = paste0(p25, "-", p75),
    `Departures before age 50 (%)` = ifelse(is.na(pct_departures_before_50), "-",
                                            sprintf("%.1f", pct_departures_before_50)),
    `Median age, graduation-year basis` = median_age_gradyear
  )
kable(s3v, caption = "**Appendix Table S3.** General obstetrics and gynecology comparator versus the three subspecialty cohorts, using the crude Open-Payments-inclusive detection-layer rate (not the primary anchored, age-stratified classifier) on the full certified cohort with Healthgrades ages. Counts and median ages therefore differ from the primary Table 3 and are a cross-cohort benchmark only.")
```

*Crude (non-age-stratified) annual departure rates are for cross-cohort
comparison only and are not the age-stratified rates used in the primary model.
Age at departure is Healthgrades profile age for GO, URPS, and MIGS and the
graduation-year proxy for the general cohort; the final column places all four on
a common graduation-year basis. URPS, Urogynecology and Reconstructive Pelvic
Surgery; GO, Gynecologic Oncology; MIGS, Minimally Invasive Gynecologic Surgery.*

Three findings follow. First, young administratively identified departures are not
unique to MIGS: the large general obstetrics and gynecology cohort also departs at
a young median age with a majority of departures before age 50, indicating that the
young-departure pattern is a property of the billing-cessation estimand rather than
a MIGS-specific defect. This is best read as a construct-validity **warning**: the
estimand captures payer exit, employer change, temporary absence, or loss of Medicare
visibility in addition to true departure from national clinical practice, and the
comparator demonstrates that the same signal produces implausibly young exits in a
generalist population. Second, the Gynecologic Oncology and URPS median departure
ages (in the sixth decade on this Healthgrades-aged benchmark, with roughly 40% of
events before age 50) are older than MIGS but are not convincingly retirement-like,
and a large share before age 50 is itself a caution rather than reassurance; we do
not treat these ages as validation of the classifier. Third, MIGS remains the
youngest-departing cohort, consistent with its being the newest subspecialty and its
weaker, voluntary-designation ascertainment, reinforcing that MIGS is the least
certain of the three.

# Appendix S3. Temporal Back-Test

To test the whole model rather than its components, we froze the active cohort at
an earlier year and rolled it forward with the age-band departure-hazard model,
comparing the projected number still active with the observed number still active
(from each physician's classified departure year). This is a closed-cohort survival
test with no graduate inflow, so it isolates the departure process. In the
calibration test the 2016 cohort was aged forward with the full-window hazards; in
the out-of-sample test the hazards were trained on 2016-2019 only and the 2019
cohort was projected to 2020-2021 (departures are too sparse before 2019 to train
on an earlier window).

```{r table-s4}
s4 <- readr::read_csv(here::here("cliff","data","temporal_backtest.csv"),
                      show_col_types = FALSE)
s4v <- s4 %>%
  transmute(
    Test = ifelse(mode == "calibration_2016_2021haz", "Calibration (2016 cohort)",
                  "Out-of-sample (2019 cohort, 2016-2019 hazards)"),
    Subspecialty = subspecialty_abbrev,
    `Base` = base, `Target` = target, `N at base` = n_base,
    `Observed active` = observed_active,
    `Predicted active` = round(predicted_active),
    `Error (%)` = sprintf("%+.1f", pct_error)
  )
kable(s4v, caption = "**Appendix Table S4.** Temporal back-test: projected versus observed active headcount for a cohort frozen at an earlier year.")
```

*Predicted counts come from aging the frozen cohort forward under the age-band
hazard model. The model reproduces the observed active trajectory to within about
1% at every horizon (mean absolute error 0.5% in the calibration test and 0.8%
out-of-sample), with a small positive bias at the horizon end reflecting a secular
rise in attrition that the earlier-window hazards under-capture, which is why the
primary analysis uses the most recent fully observable hazards. URPS, Urogynecology
and Reconstructive Pelvic Surgery; GO, Gynecologic Oncology; MIGS, Minimally
Invasive Gynecologic Surgery.*

# Appendix S4. Departure-Detection Algorithm

**Signals, priority, and weights.** Six administrative signals were integrated
(Appendix Table S2). In descending priority: (1) a lapse in ABMS Maintenance of
Certification, defined as certification lapsed with more than ten years since
certification (weight 0.95); (2) permanent NPPES deactivation without reactivation
(0.90); (3) cessation of Open Payments industry payments (0.70); (4) disenrollment
from Medicare provider enrollment in PECOS (0.80); (5) cessation of Medicare Part D
prescribing (0.60); and (6) cessation of Medicare Part B claims (0.60).

**Tie-breaking.** A higher-priority signal overrode a lower-priority one when they
disagreed; when no override applied, a weighted-consensus score (the sum of firing
signals' weights) determined status. The departure year was the confidence-weighted
consensus year across firing signals.

**Prior-positive-evidence requirement.** Each cessation signal fires only for
physicians with prior positive evidence in that source: Open Payments cessation is
evaluated only for physicians who previously received payments, Part D cessation
only for prior prescribers, and Part B cessation only for prior billers. A physician
who never received Open Payments, never prescribed, or never billed Medicare cannot
generate the corresponding cessation signal, so absence of a source is never counted
against a physician; such physicians are classified from the remaining sources.

**Non-time-limited certification and missing data.** Physicians holding lifetime
(non-time-limited) board certification cannot generate an ABMS lapse signal and are
classified from the other sources. Missing data in any source is treated as
non-informative (no signal), not as evidence of departure.

**Reactivation and re-entry.** Physicians with renewed activity after an apparent
gap (including a Medicaid-roster active override) are reclassified as active and
their departure year is withdrawn.

**Death ascertainment.** Death was captured only where a source recorded it and was
under-ascertained (the licensure-concordance analysis identified board-confirmed
deaths that the classifier did not flag). A death that is correctly captured is simply
one kind of departure and is already counted in the hazard. A *missed* death, however,
is not neutral: the physician remains in the active baseline and contributes no
departure event, which biases the estimated departure rate downward and the
completion-to-departure ratio upward. The board-confirmed deaths found during licensure review
should therefore be folded into the classifier and the effect quantified; because the
number identified was small relative to the 37 and 35 anchored departures, we expect
the direction of the headline finding to be unchanged, but this correction is
outstanding and is listed among the study limitations.

**Leave-one-source-out.** Table S5 describes the departure-*detection* layer (Step 1,
Open-Payments-inclusive) and shows how many detected departures each source supports.
In this cohort detection is carried almost entirely by the claims and payments signals:
Open Payments cessation supports about 80% of detected departures (and is the sole
signal for roughly 30%), followed by Part D and Part B; the ABMS and NPPES signals
fired for no departures, because board-certified subspecialists rarely let certification
lapse or deactivate their NPI. The heavy reliance on Open Payments at the detection
stage is exactly why the primary definition adds the Step 2 non-Open-Payments anchor:
removing the Open-Payments-sole departures reduces the detected 56 Gynecologic Oncology
and 41 URPS departures (2016-2021) to the 37 and 35 primary events, and this anchoring
motivates the licensure-concordance and back-testing analyses.

```{r table-s6}
s6 <- readr::read_csv(here::here("cliff","data","loso_source_contribution.csv"),
                      show_col_types = FALSE)
s6v <- s6 %>% transmute(
  Source = source,
  `Departures supported (n)` = n_departures_supported,
  `% of departures` = sprintf("%.1f", pct_of_departures),
  `Lost if source removed (n)` = sole_support_lost_if_removed,
  `% lost` = sprintf("%.1f", pct_lost_if_removed))
kable(s6v, caption = "**Appendix Table S5.** Leave-one-source-out: contribution of each departure signal.")
```

*Based on per-physician signal flags (credentials.retirement_signals_pivot) for
classified departures; a departure is "lost" if the removed source was its only
firing signal.*

## Exact classification rule (pseudocode)

The primary classifier is fully specified by the following Boolean rule. There is no
summed-weight threshold; the confidence weights are used only to pick the consensus
departure year among firing signals, not to decide departure status.

```text
# Per-source firing requires PRIOR POSITIVE EVIDENCE in that source
# (e.g., Part B cessation fires only for a physician who previously billed Part B).
detection_signals = {ABMS_lapse, NPPES_deactivation, PartB_cessation,
                     PartD_cessation, PECOS_disenrollment, OpenPayments_cessation}
anchor_signals    = {ABMS_lapse, NPPES_deactivation, PartB_cessation, PartD_cessation}
                    # NOTE: PECOS and Open Payments are NOT anchors

active_override    = recent Part B / Part D / PECOS activity OR Medicaid-roster
                     reactivation after the candidate year

# --- Step 1: detect a candidate departure and its year ---
candidate_departure = ( any(s fires for this physician  for s in detection_signals) )
                      AND ( NOT active_override )
candidate_year      = confidence-weighted consensus year across the firing signals
                      # weights (Table S2) resolve disagreement on WHEN, not WHETHER

# --- Step 2: primary non-Open-Payments anchor ---
primary_departure   = candidate_departure
                      AND ( any(s fires  for s in anchor_signals) )
                      # a candidate supported only by Open Payments and/or PECOS
                      # is reclassified as NOT departed for the primary analysis
```

A uniform three-year cessation rule (Part B/Part D) and the fully observable
2016-2021 window apply to both cohorts. Table S22 and the worked examples below
instantiate this rule.

## Classification decision table

The primary departure classifier is a two-step rule, not a single weighted score.
Step 1 detects a candidate departure and its year from the six administrative
signals of Appendix S4 (Table S2): each source fires only for physicians with prior
positive evidence in that source, a higher-priority signal overrides a lower-priority
one on conflict, and the departure year is the confidence-weighted consensus year
across firing signals. Step 2 is the primary-definition anchor: a candidate departure
is retained only if at least one **non-Open-Payments** signal fired, namely Medicare
Part B cessation, Medicare Part D cessation, NPPES deactivation, or ABMS certification
lapse. Candidate departures supported only by Open Payments (or only by Open Payments
and PECOS) are reclassified as not-departed for the primary analysis, because stopping
industry payments is not a validated indicator of leaving practice. An active
counter-signal at any point (recent billing, prescribing, enrollment, or a
Medicaid-roster reactivation) resets the physician to active and withdraws the
departure year. There is no numeric confidence threshold and no age or ABOG/GOBA
status weight; the decision is the Step 1 consensus gated by the Step 2 anchor.

```{r table-s22}
dt <- data.frame(
  `Departure signal` = c("ABMS MOC lapse","NPPES deactivation","Medicare Part B cessation",
             "Medicare Part D cessation","PECOS disenrollment","Open Payments cessation"),
  `Detection priority` = c(1,2,6,5,4,3),
  `Counts as non-Open-Payments anchor?` = c("Yes","Yes","Yes","Yes","No","No"),
  `Role in primary definition` = c(
    "Anchors a departure on its own","Anchors a departure on its own",
    "Anchors a departure on its own","Anchors a departure on its own",
    "Detection signal only; does not anchor","Supporting only; cannot classify a departure alone"),
  check.names = FALSE)
kable(dt, caption = "**Appendix Table S22.** Primary departure-classification decision table (the exact rule that generated every primary result). A physician is classified as departed when the Step 1 administrative-consensus rule detects a departure AND at least one non-Open-Payments anchor signal fired; departures supported only by Open Payments (with or without PECOS) are reclassified as not-departed. Any active counter-signal overrides to active. A uniform three-year cessation rule applies to both cohorts (see Right-censoring confirmability); there is no subspecialty-specific inactivity threshold.")
```

**Worked examples.** (a) A gynecologic oncologist whose Part B and Part D claims
both cease in 2019 with no later reactivation: a non-Open-Payments anchor fires, so
she is classified as departed in 2019. (b) A urogynecologist whose Open Payments
receipts stop in 2018 but who continues to bill Medicare Part B: the active
counter-signal keeps her active, and even without it no non-Open-Payments anchor
fired, so she is not a primary departure. (c) A physician with a permanent NPPES
deactivation in 2020: the deactivation anchor fires and classifies a 2020 departure.
(d) A physician whose only signals are an Open Payments cessation and a PECOS
disenrollment in 2017: neither counts as a non-Open-Payments anchor, so she is not a
primary departure, although she would have been counted under the Open-Payments-inclusive
detection layer. This is precisely why the Open-Payments-inclusive detection layer
flags 56 Gynecologic Oncology and 41 URPS departures over 2016-2021 while the primary
non-Open-Payments-anchored definition retains 37 and 35, respectively.

**The baseline gate and the departure-event definition are related but not
identical.** The 2025 active-workforce baseline (1,052 Gynecologic Oncology; 1,031
ABOG plus 264 ABU for URPS) is set by the cohorting active-status rule, a
hierarchical priority classifier (`R/retirement_sql_builder.R`) that assigns a
retirement year from, in priority order, ABMS lapse, NPPES deactivation, Open
Payments cessation, PECOS disenrollment, and Part D / Part B cessation, each under a
prior-positive-evidence and reactivation-override guard. This gate is applied once to
fix who is in the active stock; the non-Open-Payments anchor is then applied only to
the historical departure *events* that enter the hazard life table. Because the
anchor operates on the event set and not on the stock, changing it changes the
estimated hazard but does not add or remove anyone from the 2025 baseline; no
physician is "restored."

We note explicitly that the two definitions are **not** identical: the cohorting gate
can remove a physician from the 2025 active stock on Open-Payments-only inactivity
(Open Payments is a priority-3 signal in that gate), whereas the primary hazard event
set excludes Open-Payments-only departures. Both differences push the
completion-to-departure ratio in the same (upward) direction, because an
Open-Payments-inclusive stock gate removes *more* physicians (a smaller, and thus
faster-depleting, baseline) while the anchored event set counts *fewer* departures.
A fully consistent-definition sensitivity, regenerating the 2025 active stock under
the non-Open-Payments-anchored rule so that stock and flow use one identical
definition, requires re-running the cohorting classifier against the per-signal
retirement table and is an acknowledged outstanding item (it does not change the
qualitative conclusion, since both current definitions already sit far above
replacement, but it would remove the residual stock-flow definitional asymmetry).

## Right-censoring confirmability

A departure is fully observable only if the required consecutive inactive years fall within
the available data. A single uniform three-year cessation rule applies to both cohorts.
With Medicare data through 2023, a departure year D requires inactivity in D, D+1, D+2
(observable if D+2 <= 2023, i.e., D <= 2021), which defines the fully-observable 2016-2021
window for both Gynecologic Oncology and URPS; later candidate years are right-censored and
excluded from hazard estimation. There is no separate four-year rule for Gynecologic Oncology.

```{r table-s23}
cf <- data.frame(
  `Candidate departure year` = c(2019,2020,2021,2022,2023),
  `Inactive years required` = c("2019-2021","2020-2022","2021-2023","2022-2024","2023-2025"),
  `Last year needed` = c(2021,2022,2023,2024,2025),
  `Last available data year` = c(2023,2023,2023,2023,2023),
  `Fully confirmable` = c("Yes","Yes","Yes","No","No"),
  check.names = FALSE)
kable(cf, caption = "**Appendix Table S23.** Right-censoring confirmability by candidate departure year (three-year cessation rule, Medicare data through 2023). Departures through 2021 are fully observable; 2022-2023 are right-censored and excluded from hazard estimation.")
```

## Reproducibility audit table

To make every reported ratio traceable from raw events to projection, the table below
passes each departure definition through the **one** authoritative model used for the
headline result: a single pooled age-band departure hazard (estimated across the two
board-certified cohorts, gynecologic oncology and both-pathway urogynecology, to borrow
strength over the sparse per-band events; MIGS excluded), applied by dynamic aging of
each cohort's own active-age distribution with annual graduate entry. The primary
non-Open-Payments-anchored row reproduces the consolidated projection table exactly
(Gynecologic Oncology 7.35, URPS 5.83). The static single-year rate is shown alongside
the dynamic four-year departures to make the difference between the two explicit; earlier
drafts reported some sensitivities on a static, subspecialty-specific hazard, which is
the source of any small numerical differences a reader may recall.

```{r table-s4b-audit}
au <- readr::read_csv(here::here("cliff","data","departure_audit_table.csv"),
                      show_col_types = FALSE)
auv <- au %>% transmute(
  Definition = definition,
  Subspecialty = subspecialty_abbrev,
  Window = window,
  `Eligible departures (n)` = eligible_departures,
  `Static 2025 rate (%)` = sprintf("%.2f", static_2025_rate_pct),
  `Dynamic 4-yr departures` = sprintf("%.1f", dynamic_4yr_departures),
  `Dynamic annual rate (%)` = sprintf("%.2f", dynamic_annual_rate_pct),
  `Graduates/yr` = graduate_input,
  `Completion-to-departure ratio` = sprintf("%.2f", completion_to_departure_ratio),
  `Projected 2029` = sprintf("%.0f", projected_2029))
kable(auv, caption = "**Appendix Table S4b.** Reproducibility audit: every departure definition through the one authoritative pooled dynamic model. The primary (non-Open-Payments anchored) row reproduces the consolidated projection. Both the primary and the Open-Payments-inclusive definitions remain above replacement, so the qualitative finding does not depend on the source-inclusion choice.")
```

# Appendix S5. Estimated-Age Validation

The certification-anchored estimated age (reference year minus certification year
plus 30) was validated against two external sources: Healthgrades profile age and
Medicare Doctors and Clinicians graduation year (graduation year plus 27).

```{r table-s8}
s8 <- readr::read_csv(here::here("cliff","data","age_proxy_validation.csv"),
                      show_col_types = FALSE)
s8v <- s8 %>% transmute(
  Reference = reference, `N` = format(n, big.mark=",", trim=TRUE),
  `Mean error, y` = sprintf("%+.1f", mean_error),
  `SD, y` = sprintf("%.1f", sd_error),
  `Median abs. error, y (IQR)` = sprintf("%.0f (%.0f-%.0f)", median_abs_error, p25_abs, p75_abs),
  `Correlation` = sprintf("%.2f", correlation),
  `Same 5-y band (%)` = sprintf("%.1f", pct_same_band))
kable(s8v, caption = "**Appendix Table S6.** Certification-anchored estimated age versus external age sources.")
```

*Error is estimated minus reference age. The estimate correlates strongly with both
references and has a median absolute error of two years, with a small systematic
under-estimate; because the same estimate sets both the hazard-estimation and the
active-age distributions, the offset largely cancels in the departure rate.*

## Age-shift sensitivity

Because age-band hazards rise nonlinearly, a systematic age under-estimate could move
physicians across band boundaries even if the offset otherwise cancels. We recomputed the
departure rate and completion-to-departure ratio with all ages shifted +0, +2, and +4 years,
applied to both the pooled age-band hazard life table and the active-age distribution and run
through the same pooled dynamic model as the headline result (primary non-Open-Payments-anchored
definition). The +0 row therefore reproduces the Table 2 ratios exactly; the +2 and +4 rows
show the effect of the estimated-age under-estimate.

```{r table-s24}
ag <- readr::read_csv(here::here("cliff","data","age_shift_sensitivity.csv"), show_col_types = FALSE)
agv <- ag %>% dplyr::transmute(
  Cohort = subspecialty_abbrev,
  `Age shift (yr)` = sprintf("+%d", age_shift_years),
  `Departure rate (%/yr)` = sprintf("%.2f", departure_rate_pct),
  `Mean annual departures` = sprintf("%.1f", avg_annual_departures),
  `Completion-to-departure ratio` = sprintf("%.2f", replacement_ratio))
kable(agv, caption = "**Appendix Table S24.** Age-shift sensitivity. Shifting all ages up by 2-4 years modestly raises the departure rate and lowers the ratio, but both cohorts remain well above replacement, so the conclusion is robust to the estimated-age under-estimate.")
```

# Appendix S6. Cohort Derivation Flow

```{r table-s7}
s7 <- readr::read_csv(here::here("cliff","data","consort_cohort_flow.csv"),
                      show_col_types = FALSE)
lab <- c(URPS="Urogynecology and Reconstructive Pelvic Surgery",
         GO="Gynecologic Oncology", MIGS="Minimally Invasive Gynecologic Surgery")
s7 <- s7[match(c("GO","URPS","MIGS"), s7$ab), ]
s7v <- data.frame(
  Stage = c("1. Certified/designated, NPI-matched, CONUS practice",
            "2. Removed: classified deceased",
            "3. Removed: classified inactive (departed)",
            "4. ABOG-pathway active baseline",
            "5a. ABU urology-pathway net-new active identified (URPS)",
            "5b. Excluded: no datable certification year",
            "5c. ABU included in the age-structured model",
            "6. Final active baseline"),
  `Gynecologic Oncology` = c(s7$certified_matched[1], s7$removed_deceased[1],
                             s7$removed_inactive[1], s7$active_baseline[1], 0, 0, 0, s7$active_baseline_final[1]),
  `URPS` = c(s7$certified_matched[2], s7$removed_deceased[2], s7$removed_inactive[2],
             s7$active_baseline[2], s7$abu_identified[2], -s7$abu_excluded_nocert[2],
             s7$abu_included[2], s7$active_baseline_final[2]),
  `MIGS (exploratory)` = c(s7$certified_matched[3], s7$removed_deceased[3],
                           s7$removed_inactive[3], s7$active_baseline[3], 0, 0, 0, s7$active_baseline_final[3]),
  check.names = FALSE)
kable(s7v, caption = "**Appendix Table S7.** Cohort derivation from certification/designation to the final active baseline, by cohort. For URPS the final baseline is the ABOG-pathway active baseline (1,031) plus the ABU-pathway physicians included in the model (264), i.e., 1,031 + 264 = 1,295; the 6 ABU physicians without a datable certification year are excluded from both baseline and projection.")
```

*The base table is already NPI-matched and restricted to contiguous-U.S. practice.
Classified deaths are zero here, reflecting death under-ascertainment (Appendix S4),
not the absence of deaths. Of the 270 ABU urology-pathway net-new active physicians
identified, 6 lacked a datable certification year and were excluded from both the
baseline and the projection, so 264 entered the model and 1,031 + 264 = 1,295. MIGS is
an exploratory focused-practice-designation cohort.*

# Appendix S6b. Pooled Versus Unpooled Age-Band Hazards

The primary model estimates a single **pooled** age-band departure hazard across the
two board-certified primary cohorts (gynecologic oncology and both-pathway
urogynecology) and applies it to each cohort's own active-age distribution. The
exploratory MIGS cohort is **excluded** from the primary hazard because it differs in
ascertainment, certification status, and graduate numerator; a MIGS-inclusive hazard
is shown alongside as a sensitivity. Pooling is motivated by sparse per-band events:
within a single subspecialty most age bands contain only a handful of departures
(Table S6b), so unpooled subspecialty-specific hazards are unstable. The table reports,
for each age band, the departure events and person-years by subspecialty, the unpooled
Gynecologic Oncology and URPS hazards, the pooled hazard actually used, and the
MIGS-inclusive pooled hazard as a sensitivity.

```{r table-s6b-band}
hb <- readr::read_csv(here::here("cliff","data","hazard_by_band_pooled_vs_unpooled.csv"),
                      show_col_types = FALSE)
hbv <- hb %>% dplyr::transmute(
  `Age band` = band,
  `GO events` = go_events, `GO PY` = go_py,
  `URPS events` = urps_events, `URPS PY` = urps_py,
  `MIGS events` = migs_events,
  `GO hazard (unpooled)` = sprintf("%.4f", go_hazard_unpooled),
  `URPS hazard (unpooled)` = sprintf("%.4f", urps_hazard_unpooled),
  `Pooled hazard (used, GO+URPS)` = sprintf("%.4f", pooled_hazard),
  `Pooled +MIGS (sensitivity)` = sprintf("%.4f", pooled_incl_migs_hazard))
kable(hbv, caption = "**Appendix Table S6b.** Age-band departure events, person-years, and hazards: unpooled (subspecialty-specific) versus the pooled hazard used in the primary model (GO + both-pathway URPS only, primary non-Open-Payments-anchored definition, 2016-2021), with the MIGS-inclusive pooled hazard as a sensitivity. The primary hazard life table totals 36 Gynecologic Oncology and 35 URPS events (71 pooled); adding the exploratory MIGS cohort raises the pool to 81 events. The 37th Gynecologic Oncology anchored departure is dated in its first observable year and contributes no prior at-risk person-time.")
```

```{r table-s6b-rate}
hr <- readr::read_csv(here::here("cliff","data","hazard_rate_pooled_vs_unpooled.csv"),
                      show_col_types = FALSE)
hrv <- hr %>% dplyr::transmute(
  Subspecialty = subspecialty_abbrev,
  `Weighted 2025 rate, pooled GO+URPS hazard (%)` = sprintf("%.2f", rate_pooled_pct),
  `Weighted 2025 rate, unpooled hazard (%)` = sprintf("%.2f", rate_unpooled_pct),
  `Weighted 2025 rate, pooled +MIGS sensitivity (%)` = sprintf("%.2f", rate_pooled_incl_migs_pct))
kable(hrv, caption = "**Appendix Table S6c.** Weighted 2025 static departure rate under the pooled hazard (used) versus each subspecialty's own unpooled hazard. Under unpooled hazards URPS (1.01%) exceeds Gynecologic Oncology (0.90%); under the pooled hazard applied to each age distribution the ordering reverses (Gynecologic Oncology 0.96%, URPS 0.83%). The pooling assumption, not separate estimation, therefore determines which subspecialty's rate is higher, so the per-subspecialty rates should be read as one pooled hazard filtered through two age structures, not as independently estimated specialty hazards.")
```

Because pooling assumes a common age-conditioned hazard across subspecialties that
differ in practice structure and Medicare visibility, we present it transparently
rather than as specialty-specific estimation. A hierarchical (partially pooled)
survival model would preserve subspecialty identity while borrowing strength across
the sparse bands and is the preferred next step; the fully pooled and fully unpooled
rates above bracket its result, and both leave every cohort far above replacement.

# Appendix S7. Individual-Level Adjudication Protocol

The licensure-concordance analysis (main text) compares the classifier with an
independent administrative reference but is not an individual-level validation. We
therefore specify a physician-level manual adjudication over a single definitive
sample (`cliff/data/classifier_adjudication_sample.csv`; scripts/classifier_adjudication.R):
**all** classified clinical-practice departures in Gynecologic Oncology and URPS
(2016-2021), enriched with all departures under age 50 and all departures lacking
independent corroboration, plus a control set of physicians classified continuously
active (to detect false negatives, the error that most threatens the replacement
result). For each physician a scoring sheet records the classifier label, classified
departure year, estimated age, and firing signals, with blank fields for two
independent reviewers to record an adjudicated active/departed status, a departure
year, and a confidence rating, using public-record verification (state licensure,
institutional affiliation, recent publications or presentations, and obituary or
death-index searches). Positive and negative predictive values, sensitivity,
specificity, inter-reviewer agreement (Cohen kappa), and adjudication-corrected
departure rates and completion-to-departure ratios will be reported for Gynecologic
Oncology and ABOG-pathway URPS, with disagreements resolved by consensus. ABU
urology-pathway departures are not independently classifiable in this cohort (the ABU
roster carries no longitudinal activity data), so ABU-specific predictive values and
corrected hazards are not adjudicated; the projection borrows the ABOG-URPS hazard for
the ABU component. The sampling frame and instrument are prespecified and complete, and
the two-reviewer adjudication is pending; because this individual-level validation is
not yet complete, the primary departure endpoint is reported throughout as a provisional
administrative classification, and the headcount conclusion is supported instead by the
tipping-point robustness analysis (Appendix S10), which shows that overturning it would
require a large fraction of true departures to be missed. We do not present the classifier
as individually validated.

**Automated corroboration is meaningful only for the Open-Payments-inclusive
detection layer, and is tautological for the primary definition.** Table S7b below
reports, for the Open-Payments-inclusive *detection* layer (56 Gynecologic Oncology and
41 URPS candidate departures, 2016-2021), the fraction with at least one non-Open-Payments
signal (NPPES deactivation, Medicare billing cessation in the departure year or the year
after, or an ABMS lapse within two years): 36% and 46% respectively. This corroboration
is informative there only because that layer includes Open-Payments-only events that the
non-Open-Payments signals can either confirm or fail to confirm. It is **not** a
validation of the primary definition: the primary non-Open-Payments-anchored endpoint
retains exactly the 37 Gynecologic Oncology and 35 URPS candidates that carry such an
anchor, so by construction 100% of primary departures are non-Open-Payments-corroborated.
That 100% is definitional, not independent evidence, and does not establish positive
predictive value; genuine individual-level validation requires the pending chart
adjudication above.

```{r table-s7-corr}
cc <- readr::read_csv(here::here("cliff","data","classifier_corroboration.csv"), show_col_types = FALSE)
cc <- cc[match(c("GO","URPS"), cc$subspecialty_abbrev), ]
ccv <- data.frame(
  Cohort = cc$subspecialty_abbrev,
  `Classified departures (n)` = cc$n_departures,
  `NPPES deactivated (%)` = cc$pct_nppes_deactivated,
  `Medicare billing ceased (%)` = cc$pct_billing_ceased,
  `ABMS retirement concordant (%)` = cc$pct_abms_concordant,
  `Any signal (provisional PPV proxy, %)` = cc$ppv_proxy_corroborated,
  check.names = FALSE)
kable(ccv, caption = "**Appendix Table S7b.** Automated corroboration of the Open-Payments-INCLUSIVE detection-layer departures (56 Gynecologic Oncology, 41 URPS) by non-Open-Payments administrative signals. This is not a validation of the primary definition: the primary non-Open-Payments-anchored endpoint (37 and 35) is 100% non-Open-Payments-corroborated by construction. Neither is a substitute for the pending chart adjudication.")
```

# Appendix S8. Feminization Sensitivity

These subspecialties are feminizing: the female share of the cohort rose from 33%
among physicians certified through 2005 to 60%, 72%, and 71% among those certified
in 2006-2012, 2013-2018, and 2019 onward. Women leave clinical practice at a higher
hazard than men (hazard ratio ~1.4). We modeled this with a sex-stratified dynamic
projection in which the empirical age-band hazard is split into sex-specific hazards
(female = 1.4 x male, calibrated to reproduce the observed mixed-sex rate) and
graduates enter at the female share of the OB/GYN resident pool that feeds these
fellowships (about 85%, ACGME), with the observed recent-subspecialist share (72%)
as a lower bound. The URPS results here are **ABOG-pathway only**: sex is ascertainable
only for the ABOG-certified cohort (n = 1,031; the external ABU roster carries no sex
field), and the 85% female entrant share is the obstetrics-gynecology resident pool
feeding the ABOG pathway, not the urology pathway. The sex-neutral URPS ratio below is
therefore the ABOG-pathway value computed on the sex-known subsample, which is lower than
the both-pathway primary ratio (`r get_replacement_ratio("URPS")`) reported in the main text.

```{r table-s10}
s10 <- readr::read_csv(here::here("cliff","data","feminization_sensitivity.csv"),
                       show_col_types = FALSE)
s10v <- s10 %>% transmute(
  Cohort = subspecialty_abbrev,
  `Active female (%)` = sprintf("%.1f", active_female_pct),
  `Ratio, sex-neutral` = sprintf("%.2f", ratio_sexneutral),
  `Ratio, 72% female entrants` = sprintf("%.2f", ratio_fem_entrants72),
  `Ratio, 85% female entrants` = sprintf("%.2f", ratio_fem_entrants85),
  `2029 change, 85% (%)` = sprintf("%+.1f", pct_change_fem85))
kable(s10v, caption = "**Appendix Table S8.** Feminization sensitivity: sex-stratified projection versus the sex-neutral ratio (URPS is ABOG-pathway, sex-known subsample; not the both-pathway primary).")
```

*All cohorts remain above replacement. The ratios are essentially unchanged and
insensitive to the entrant female share (72% versus the ~85% resident pool), because
entrants are young and depart at low rates regardless of sex; near-term departures
come from older, male-dominated cohorts, so the early-career feminization effect
emerges only over a longer horizon as the majority-female younger cohort ages (URPS
here uses the ABOG-pathway cohort, for which sex is recorded).*

# Appendix S9. Clinical Full-Time-Equivalent Sensitivity

The primary analysis counts physicians (headcount). Because clinical effort per
physician is below one and varies by subspecialty, we report an exploratory
clinical full-time-equivalent (FTE) sensitivity. Specialty-specific clinical-effort
estimates are available but derive from surveys rather than a national FTE census:
gynecologic oncologists devoted a median 63% of professional effort to clinical
activities (2020 Society of Gynecologic Oncology State of the Society Survey),[@temkin2022fte]
and one survey of 118 female pelvic medicine and reconstructive surgery physicians
reported about 80% clinical or operative effort.[@aua2024fpmrs] We treat the URPS
factor as provisional (sensitivity range 0.60 to 0.90) because that sample may
over-represent urology-pathway practice; MIGS effort is not estimated.

```{r table-s9}
bl <- setNames(workforce_summary$baseline_2025, workforce_summary$subspecialty_abbrev)
fte <- data.frame(
  Cohort = c("Gynecologic Oncology", "Urogynecology and Reconstructive Pelvic Surgery (both-pathway)"),
  `Active headcount` = c(format(bl[["GO"]], big.mark=","), format(bl[["URPS"]], big.mark=",")),
  `Clinical-effort factor` = c("0.63", "0.80 (0.60-0.90)"),
  `Approximate clinical FTEs` = c(
    format(round(bl[["GO"]]*0.63), big.mark=","),
    sprintf("%s (%s-%s)", format(round(bl[["URPS"]]*0.80), big.mark=","),
            format(round(bl[["URPS"]]*0.60), big.mark=","), format(round(bl[["URPS"]]*0.90), big.mark=","))),
  check.names = FALSE)
kable(fte, caption = "**Appendix Table S9.** Clinical-FTE sensitivity applying specialty-specific clinical-effort factors to the active headcount.")
```

*Clinical FTE (the share of professional effort devoted to clinical care) differs
from employment FTE (hours worked relative to a full-time standard) and from
headcount; no national census resolves all three for these subspecialties.
Critically, applying the same clinical-effort factor to both the incumbent workforce
and incoming graduates rescales absolute capacity but leaves the completion-to-departure ratio
unchanged; the ratio would move only if clinical effort differed systematically
between existing, departing, and newly graduating physicians (for example, if new
graduates entered at lower effort or effort declined with age), which available data
do not resolve. These estimates are therefore a sensitivity, not a primary workforce
measure.*

## Clinical and operative effort of new graduates

The completion-to-departure ratio changes if incoming graduates deliver less direct clinical and
operative capacity per physician than the URPS and gynecologic oncology surgeons they
replace, even when total professional hours are preserved. Several converging trends
could produce such a shift: a larger share of professional time in administrative and
nonclinical roles, more part-time and reduced-hours practice, career interruptions for
childbearing in a predominantly (about 85%) female entering cohort, and lower operative
volume. Existing data are consistent with a clinical-and-operative, rather than
total-hours, gap: in obstetrics-gynecology direct patient-care hours were similar by sex,
but nonclinical hours and annual procedural or delivery volume were lower for women and
total professional effort was about 91% of men's after age adjustment.[@benedetti2004productivity; @romero2025feminization]
We therefore stress the entrant-to-incumbent clinical and operative effort as an explicit
multiplier; the effective completion-to-departure ratio is the headcount ratio times that multiplier:

```{r table-s9b}
r <- workforce_summary
eff <- data.frame(
  Cohort = r$subspecialty_abbrev,
  `Headcount ratio` = sprintf("%.2f", r$replacement_ratio),
  `x0.85` = sprintf("%.2f", r$replacement_ratio*0.85),
  `x0.75` = sprintf("%.2f", r$replacement_ratio*0.75),
  `x0.65` = sprintf("%.2f", r$replacement_ratio*0.65),
  check.names = FALSE)
kable(eff, caption = "**Appendix Table S10.** Effective completion-to-departure ratio under an entrant-to-incumbent clinical-and-operative-effort multiplier (0.85 to 0.65).")
```

*Every cohort remains above replacement even at a 0.65 multiplier, a 35% reduction in the
clinical and operative effort of new graduates relative to the surgeons they replace
(Gynecologic Oncology `r sprintf("%.1f", 0.65*as.numeric(get_replacement_ratio("GO")))`, URPS `r sprintf("%.1f", 0.65*as.numeric(get_replacement_ratio("URPS")))`, MIGS `r sprintf("%.1f", 0.65*as.numeric(get_replacement_ratio("MIGS")))`). The magnitude of any such generational
shift is uncertain, so this is presented as a stress test rather than a central estimate.
A related forward risk, not modeled here, is that younger cohorts anticipate earlier
retirement, which would raise future departure rates beyond this horizon.[@muzyka2025generational]*

# Appendix S10. Break-Even Thresholds

To characterize how much the conclusion can absorb, we computed, for each cohort, the
single-parameter value at which the completion-to-departure ratio would fall to 1.0 (the point at
which graduate output stops replacing departures), holding the other inputs at their
primary values.

```{r table-s11}
bt <- readr::read_csv(here::here("cliff","data","breakeven_thresholds.csv"),
                      show_col_types = FALSE)
btv <- bt %>% transmute(
  Cohort = subspecialty_abbrev,
  `Completion-to-departure ratio` = sprintf("%.2f", replacement_ratio),
  `Break-even departure rate` = sprintf("%.1f%% (%.1fx empirical)", breakeven_departure_rate_pct, breakeven_rate_multiple),
  `Break-even graduate drop` = sprintf("-%d%% (to %d/yr)", graduate_drop_pct, breakeven_graduates),
  `Break-even effort multiplier` = sprintf("%.2f", breakeven_effort_multiplier))
kable(btv, caption = "**Appendix Table S11.** Single-parameter values that would reduce each cohort to replacement (ratio = 1.0).")
```

```{r s10-prose, results='asis'}
cat(sprintf(
  "*No single parameter reaches replacement at a plausible value: the departure rate would have to rise to %.1f%% to %.1f%% per year (%.1f to %.1f times the empirical rate, versus the 1.5%% to 1.7%% national all-surgeon benchmark), graduate output would have to fall %d%% to %d%%, or new graduates would have to work at %d%% to %d%% of incumbent clinical effort. Only a compound worst-case, the departure rate tripled and graduate output reduced 30%% and the entrant effort multiplier at 0.65 simultaneously, pushes Gynecologic Oncology (%.2f) and URPS (%.2f) below replacement; the exploratory MIGS cohort remains above (%.2f). URPS, with the lowest ratio, is the most sensitive of the two board-certified subspecialties.*",
  min(bt$breakeven_departure_rate_pct), max(bt$breakeven_departure_rate_pct),
  min(bt$breakeven_rate_multiple), max(bt$breakeven_rate_multiple),
  min(bt$graduate_drop_pct), max(bt$graduate_drop_pct),
  as.integer(round(100*min(bt$breakeven_effort_multiplier))), as.integer(round(100*max(bt$breakeven_effort_multiplier))),
  bt$compound_worstcase_ratio[bt$subspecialty_abbrev=="GO"],
  bt$compound_worstcase_ratio[bt$subspecialty_abbrev=="URPS"],
  bt$compound_worstcase_ratio[bt$subspecialty_abbrev=="MIGS"]))
```

# Appendix S11. Operative-Workforce Analysis (Exploratory)

## S11.1 Rationale and definitions

Because a surgeon can leave operative practice while remaining clinically active,
headcount replacement (the primary outcome) need not equal replacement of operative
capacity. Retirement from operative practice is a recognized workforce transition,
distinct from and typically earlier than complete clinical retirement: in a population
survey of general surgeons the median age of stopping operating (65 years, interquartile
range 60-67.5) preceded stopping all clinical work (70 years) by roughly five years.[@poushay2018; @conn2018; @greenfield1994; @acog2018latecareer]
Prior evidence is almost entirely survey-based, whereas national administrative attrition
studies define workforce exit through cessation of evaluation-and-management claims or
clinical payments rather than of operations.[@elemosho2026; @hedden2017] We therefore
conducted an exploratory analysis operationalizing operative-workforce exit longitudinally
from Medicare Part B by-service claims (procedure-level HCPCS codes per physician per year),
restricted to Gynecologic Oncology and URPS; MIGS was excluded because only about 13% of
that cohort was visible as Medicare operators. **The URPS operative analysis uses the
ABOG-pathway URPS cohort only** (active n = 1,031), because the ABU (urology-pathway)
physicians are not claims-linkable in this cohort; both the operator denominator and the
graduate-entry numerator are therefore ABOG-pathway. All URPS operative figures below should
be read as ABOG-pathway URPS, not the both-pathway workforce (1,295) used in the primary
headcount analysis. As a share of the both-pathway workforce the visible-operator percentage
would be correspondingly lower (e.g., 613/1,295 = 47% rather than 613/1,031 = 60%).

The qualifying major-operation HCPCS list is the subspecialty operative code set from the
project procedure reference (58210-58958, 38570-49255, 56630-58240 and 57xxx pelvic-floor
reconstruction, 64xxx neuromodulation, and 58541-58573 laparoscopic hysterectomy families).
An **established operator** performed at least three qualifying operations per year in at
least two of the three preceding years, preventing an isolated procedure from counting as
mature operating. The measured event was **sustained cessation of Medicare-observed
qualifying surgery**: an established operator whose final qualifying operation was followed
by three consecutive zero-operation years with no later resumption. With Medicare data
ending 2023, the event is datable only through 2020, so events were counted over the fully
observable window 2016-2020. We deliberately use "sustained cessation of Medicare-observed
qualifying surgery" rather than "operative retirement" for the administrative event,
because (as S11.4 shows) the two are not equivalent.

## S11.2 Corrected denominator, cessation, and volume

Operator prevalence is expressed as a share of the **active** board-certified cohort
(is_retired_for_cohorting = FALSE), not the total cohort including retirees; the earlier
denominator wrongly included retired physicians and understated the share.

```{r table-s12}
ow <- readr::read_csv(here::here("cliff","data","operative_workforce.csv"),
                      show_col_types = FALSE)
ow <- ow[match(c("GO","URPS"), ow$subspecialty_abbrev), ]
owv <- data.frame(
  Cohort = ow$subspecialty_abbrev,
  `Active cohort (n)` = ow$n_active_cohort,
  `Established Medicare operators (n, % of active)` =
    sprintf("%d (%.0f%%)", ow$medicare_visible_operators, ow$pct_of_active_visible),
  `Sustained-cessation rate (%/yr)` = sprintf("%.1f", ow$operative_cessation_rate_pct),
  `Still billing clinically after last operation (%)` = ow$pct_still_billing_after,
  check.names = FALSE)
kable(owv, caption = "**Appendix Table S12.** Established Medicare operators and sustained cessation of Medicare-observed qualifying surgery, active board-certified cohort (2016-2020, Medicare-provisional).")
```

## S11.3 Time-to-event operative entry and the follow-up-time artifact

A fixed graduate ramp-up window makes the operator-entry numerator depend mechanically on
follow-up length: new surgeons accrue a Medicare surgical panel only gradually, so a short
window counts few of them. Estimating the cumulative incidence of becoming an established
Medicare operator with identical follow-up per certification cohort shows this directly,
the conversion probability roughly triples between 3 and 8 years. The earlier fixed-ramp
"matched ratio" (as low as 0.34 at a 3-year window) was therefore a follow-up-time
artifact, not evidence of a capacity deficit.

```{r table-s13}
tt <- readr::read_csv(here::here("cliff","data","operative_tte.csv"), show_col_types = FALSE)
tt <- tt[match(c("GO","URPS"), tt$subspecialty_abbrev), ]
ttv <- data.frame(
  Cohort = tt$subspecialty_abbrev,
  `Certification cohort (2013-2020, n)` = tt$cohort_grads_2013_2020,
  `Established operator by 3 y (%)` = sprintf("%.0f", 100*tt$ci_operator_by3yr),
  `by 5 y (%)` = sprintf("%.0f", 100*tt$ci_operator_by5yr),
  `by 8 y (%)` = sprintf("%.0f", 100*tt$ci_operator_by8yr),
  check.names = FALSE)
kable(ttv, caption = "**Appendix Table S13.** Cumulative incidence of becoming an established Medicare operator by years since certification (time-to-event, identical follow-up per cohort). The steep rise from 3 to 8 years is the follow-up-time artifact underlying the earlier fixed-ramp estimate.")
```

```{r table-s14}
cp <- readr::read_csv(here::here("cliff","data","operative_capacity.csv"), show_col_types = FALSE)
cp <- cp[match(c("GO","URPS"), cp$subspecialty_abbrev), ]
cpv <- data.frame(
  Cohort = cp$subspecialty_abbrev,
  `Operators produced/yr (grads x 8-y conversion)` = sprintf("%.1f", cp$operators_produced_per_yr),
  `Operators ceasing/yr` = sprintf("%.1f", cp$retirers_per_yr),
  `Operator-count replacement` = sprintf("%.2f", cp$operator_count_replacement),
  `Entrant vs retiring per-capita op volume` = sprintf("%.1f vs %.1f", cp$entrant_mean_annual_op_volume_yr3to5, cp$retiring_mean_annual_op_volume),
  `Volume-weighted capacity ratio` = sprintf("%.2f", cp$operative_workforce_capacity_ratio),
  check.names = FALSE)
kable(cpv, caption = "**Appendix Table S14.** Operator-count and procedure-volume-weighted operative-workforce replacement. Volume unit = annual qualifying operative services. The volume ratio compares rising entrants (post-certification years 3-5) against tapering retirees (final two operative years), which biases it upward; both estimates are above one but are NOT interpreted given the endpoint's poor validity (S11.4).")
```

## S11.4 Concordance with complete departure: the validity finding

Each apparent operative cessation was compared for concordance with independent signals
of *complete clinical-practice departure*: whether total Medicare billing (across all
HCPCS, not only operations) also stopped, and whether the NPI was deactivated in NPPES. A
cessation is concordant with complete departure only if at least one fired. These signals
detect departure from clinical medicine, not cessation of operating specifically, so they
cannot confirm true operative cessation, by the analysis's own premise a physician can stop
operating while continuing clinic billing with an active NPI. The concordance is therefore
interpreted only as evidence that most operative-cessation events are not complete
departures, not as a positive predictive value for operative cessation.

```{r table-s15}
aj <- readr::read_csv(here::here("cliff","data","operative_adjudication.csv"), show_col_types = FALSE)
aj <- aj[match(c("GO","URPS"), aj$subspecialty_abbrev), ]
ajv <- data.frame(
  Cohort = aj$subspecialty_abbrev,
  `Apparent cessations (n)` = aj$n_cessations,
  `Total billing continued (%)` = aj$pct_total_billing_continued,
  `Total billing stopped (%)` = aj$pct_total_billing_stopped,
  `NPPES deactivated (%)` = aj$pct_nppes_deactivated,
  `Age <50 at event (%)` = aj$pct_age_under_50,
  `Concordant with complete departure (%)` = aj$pct_corroborated_true_exit,
  check.names = FALSE)
kable(ajv, caption = "**Appendix Table S15.** Concordance of apparent operative cessations with independent signals of complete clinical-practice departure. Only a small minority coincide with a complete-departure signal, and most physicians continued billing clinically; because a true operative retiree may keep billing, this does not establish whether events were true operative cessation or loss of Medicare visibility.")
```

Only `r get_op_corroborated("GO")`% (Gynecologic Oncology) and `r get_op_corroborated("URPS")`%
(URPS) of apparent operative cessations coincided with a complete-departure signal;
`r get_op_still_billing("GO")`% and `r get_op_still_billing("URPS")`% continued billing
clinically. Because Medicare fee-for-service cannot distinguish true cessation of surgery
from loss of claims visibility, the corrected above-one replacement estimates in S11.3 are
reported but not interpreted. Reliable measurement of
operative-capacity replacement requires all-payer claims or institutional operative-log
data capturing the commercial and facility-billed surgery Medicare omits. A stratified
sample of apparent operative cessations selected for adjudication (all events younger than 50,
all with continued billing, and corroborated exits) is provided in the reproducibility archive
(`cliff/data/operative_cessation_adjudication_sample.csv`).

## S11.5 URPS pathway (ABU) hazard sensitivity

The both-pathway URPS baseline assigns ABU-certified urogynecologists the ABOG-URPS
departure hazard because ABU-pathway hazards are not independently estimable in this
cohort. The table brackets that assumption by modeling the ABU component separately and
scaling its hazard from half to twice the ABOG rate (approximate deterministic projection;
the authoritative both-pathway ratio is the Monte Carlo value in the main table).

```{r table-s16}
ab <- readr::read_csv(here::here("cliff","data","abu_pathway_sensitivity.csv"), show_col_types = FALSE)
kable(ab %>% dplyr::transmute(
  Scenario = scenario, `Baseline (n)` = baseline, `Annual grads` = annual_grads,
  `Avg annual departures` = sprintf("%.1f", avg_annual_departures),
  `Completion-to-departure ratio` = sprintf("%.2f", replacement_ratio)),
  caption = sprintf("**Appendix Table S16.** URPS completion-to-departure ratio under separate ABU-pathway modeling and ABU-hazard scaling. URPS remains well above replacement across all scenarios. The both-pathway (ABU = ABOG hazard) value here (%.2f) is computed with a deterministic baseline-rate approximation and is therefore not expected to match the primary both-pathway ratio (%s), which comes from the age-structured dynamic model with year-by-year aging into higher-hazard bands; the two differ only by the deterministic-versus-dynamic method, not by inputs.",
    ab$replacement_ratio[ab$scenario=="Both-pathway (ABU=ABOG hazard)"], get_replacement_ratio("URPS")))
```

# Appendix S12. Multidimensional Sensitivity Grid

The main-text sensitivity analyses are one-way (single-lever) deviations from the primary
projection, and all leave both board-certified cohorts above replacement. To report the
fully-crossed grid honestly, we crossed the observation window (fully-observable 2016-2021,
drop-2 2016-2019, and the provisional full 2016-2023, which may be upward biased), a departure-rate multiplier
(1x/2x/3x), and graduate practice-conversion (1.00/0.85/0.70), giving 27 cells per cohort;
each cell's ratio is the window ratio scaled by conversion divided by the rate multiplier.

```{r table-s17}
gs <- readr::read_csv(here::here("cliff","data","sensitivity_grid_summary.csv"), show_col_types = FALSE)
gs <- gs[match(c("GO","URPS","MIGS"), gs$subspecialty_abbrev), ]
gsv <- data.frame(
  Cohort = gs$subspecialty_abbrev,
  `Cells` = gs$n_cells,
  `Above replacement (>1.05)` = gs$n_above,
  `At replacement (0.95-1.05)` = gs$n_at,
  `Below replacement (<0.95)` = gs$n_below,
  `Worst cell` = sprintf("%.2f", gs$worst_ratio),
  `One-way minimum` = sprintf("%.2f", gs$oneway_min),
  check.names = FALSE)
kable(gsv, caption = "**Appendix Table S17.** Multidimensional sensitivity grid (27 cells per cohort). Every one-way sensitivity stays above replacement; the only at-replacement (Gynecologic Oncology) and below-replacement (URPS) cells occur only where the provisional full window, a tripled departure rate, and 70% graduate conversion coincide.")
```

*All one-way minima exceed 1.05. At the joint extreme (full window, 3x rate, 0.70 conversion)
Gynecologic Oncology was at replacement (1.03) and URPS was below replacement (0.88), a compound
worst case rather than a plausible single deviation; the exploratory MIGS cohort stays above
replacement in every cell.*

# Appendix S13. MIGS Exploratory Focused-Practice Cohort

Minimally invasive gynecologic surgery (MIGS) is analyzed separately from the two
board-certified subspecialties because it differs fundamentally: it is a voluntary
ABOG focused-practice *designation* rather than a board certification, its stock is
designation-based, its inflow is NRMP/AAGL match fills rather than ACGME completers,
training non-completion is unobserved, and it is not governed by the same stock-flow
definitions. It is therefore excluded from the main analysis and reported here only
as an exploratory cohort.

```{r table-s18}
kable(data.frame(
  Quantity = c("2025 baseline (n)", "Annual departure rate (%)", "Mean annual departures",
               "Annual entrants (NRMP/AAGL match fill)", "Completion-to-departure ratio",
               "Projected 2029 (n)", "Estimated median departure age (y)"),
  Value = c(get_baseline("MIGS"), get_annual_rate("MIGS"), get_avg_retirements("MIGS"),
            as.character(get_annual_entrants("MIGS")), get_replacement_ratio("MIGS"),
            get_projected("MIGS"),
            as.character(readr::read_csv(here::here("cliff","data","age_at_departure_by_subspecialty.csv"),
                         show_col_types=FALSE) |> dplyr::filter(subspecialty_abbrev=="MIGS") |> dplyr::pull(median_age))),
  check.names = FALSE),
  caption = "**Appendix Table S18.** MIGS exploratory focused-practice cohort (designation-based; NRMP/AAGL inflow). Reported for completeness, not pooled with the board-certified subspecialties.")
```

*The estimated median classified departure age of 34 years is an important validity
warning: it may reflect nonretirement workforce exits, but may equally reflect
transient administrative invisibility, designation uptake, young physicians without
Medicare panels, relocation, or coding and cohort-definition artifacts. The general
obstetrics-gynecology comparator (Appendix S2) shows young administrative departures
across the whole specialty rather than uniquely in MIGS, but does not validate the
individual MIGS events. The MIGS replacement estimate should not be interpreted as a
measure of the national MIGS clinical-practice workforce.*

# Appendix S14. Certification-to-First-Medicare-Activity Transition

The primary model adds fellowship completers to the active board-certified stock in
their entry year. To test that immediate-entry assumption, we estimated a certification-to-first-Medicare-activity
transition by linking board-certified GO and URPS certification cohorts (2013-2019,
with at least four years of follow-up) to their first observed Medicare activity. This
is a proxy for, not a direct measure of, the fellowship-completion-to-active-practice
lag: certification is not fellowship graduation (and can follow practice entry), and
first Medicare billing is not first active practice (non-Medicare practice is invisible). The transition is reported as a **timing ramp**
normalized to eventual activation, so it measures how quickly graduates reach observed
active practice rather than the Medicare-visibility ceiling; the separate ever-billed
ceiling (the fraction ever observed in Medicare) reflects non-Medicare practice, not
non-entry, and is not applied as a yield penalty.

```{r table-s19}
tr <- readr::read_csv(here::here("cliff","data","graduation_active_transition.csv"), show_col_types = FALSE)
trv <- tr %>% dplyr::filter(k <= 4) %>%
  dplyr::mutate(pct = sprintf("%.0f%%", 100*cum_active_fraction)) %>%
  dplyr::select(subspecialty_abbrev, k, pct) %>%
  tidyr::pivot_wider(names_from = k, values_from = pct, names_prefix = "By yr ")
cel <- tr %>% dplyr::filter(k == 0) %>% dplyr::transmute(subspecialty_abbrev,
        `Ever-billed ceiling` = sprintf("%.0f%%", 100*ever_billed_ceiling))
kable(dplyr::left_join(trv, cel, by = "subspecialty_abbrev"),
      caption = "**Appendix Table S19.** Certification-to-first-Medicare-activity transition: cumulative fraction of a certification cohort with observed Medicare activity by k years after certification (timing ramp, normalized to eventual activation, among certificants ever observed in Medicare), and the ever-billed Medicare-visibility ceiling (not a yield penalty).")
```

```{r table-s20}
pj <- readr::read_csv(here::here("cliff","data","graduation_active_transition_projection.csv"), show_col_types = FALSE)
kable(pj %>% dplyr::transmute(
  Cohort = subspecialty_abbrev,
  `Baseline 2025` = format(baseline, big.mark = ","),
  `Projected 2029 (immediate entry)` = format(projected_2029_immediate, big.mark = ","),
  `Projected 2029 (transition ramp)` = format(projected_2029_ramped, big.mark = ","),
  `Growth deferred beyond 2029` = sprintf("%d (%d%%)", difference, pct_of_growth_deferred)),
  caption = "**Appendix Table S20.** Transition-adjusted projection. Applying the empirical entry ramp defers part of the projected four-year growth beyond 2029; the workforce still grows and the completion-to-departure ratio (a flow balance) is unchanged.")
```

*Among certificants eventually observed in Medicare, 96% appeared within three years, so
observed activity accrues over roughly two to three years; only 89% (GO) and 84% (URPS)
of certificants were ever Medicare-visible, and while the remainder plausibly reflects
non-Medicare practice, some may not enter U.S. subspecialty practice. The immediate-entry
assumption thus modestly overstates near-term growth but not the direction of the conclusion. The completion-to-departure ratio compares
annual entrants with annual departures and is unaffected by entry timing; the transition
function refines the projected stock trajectory, not the replacement balance.*

# Appendix S15. Open Payments Construct-Validity Sensitivity

Cessation of Open Payments (industry-payment) reporting supported about 80% of classified
departures under the prior definition and was the sole firing source for about 30%, whereas
ABMS lapse and NPPES deactivation identified none, and Part D and Part B contributed fewer
unique cases. Because stopping industry payments is not a validated indicator of leaving
clinical practice, **the primary departure definition requires a non-Open-Payments
corroborating source** (Medicare Part B or D, NPPES deactivation, or ABMS lapse); an Open
Payments cessation alone no longer classifies a departure. The table below passes both the
Open-Payments-inclusive definition and the primary non-Open-Payments-anchored definition
through the same pooled dynamic model used for the headline result, so the primary row
reproduces Table 2 exactly. Removing the Open-Payments-only departures lowers the departure
rate and raises the ratio; we report this as the direction of the arithmetic only and do not
assert which definition is closer to truth, which awaits the individual chart adjudication.
Both cohorts remain above replacement under either definition.

```{r table-s21}
os <- readr::read_csv(here::here("cliff","data","open_payments_sensitivity.csv"), show_col_types = FALSE)
lab <- c(`op_inclusive` = "Open Payments-inclusive definition",
         `non_op_anchored (primary)` = "Non-Open-Payments anchor required (primary)")
osv <- os %>% dplyr::transmute(
  Definition = lab[rule], Cohort = subspecialty_abbrev,
  `Departure rate (%/yr)` = sprintf("%.2f", departure_rate_pct),
  `Mean annual departures` = sprintf("%.1f", avg_annual_departures),
  `Completion-to-departure ratio` = sprintf("%.2f", replacement_ratio))
kable(osv, caption = "**Appendix Table S21.** Departure rate and fellowship-completion-to-departure ratio under the Open-Payments-inclusive and primary non-Open-Payments-anchored definitions, both passed through the same pooled dynamic model (the primary row reproduces Table 2). Both board-certified cohorts remain above replacement under either definition; the direction of the difference is reported without a claim as to which definition is closer to truth.")
```

*Both cohorts remain above replacement under either definition. Requiring a
non-Open-Payments anchor (Medicare Part B or D cessation, NPPES deactivation, or ABMS
lapse; excluding Open-Payments-only departures) lowers the departure rate and raises
the ratio. We report this direction of the arithmetic only and make no claim as to
which definition is closer to truth, which awaits the individual chart adjudication;
in particular we do not assert that the Open-Payments-inclusive definition was
conservative, because that would require knowing which Open-Payments-only events are
true departures. The error that could threaten the result is missed departures (false
negatives), addressed by the tipping-point analysis and the pending active-control
adjudication (Appendix S7).*

# Age-Specific Mortality Sensitivity (Reviewer #2)

```{r table-s25}
ms <- readr::read_csv(here::here("cliff","data","mortality_sensitivity.csv"), show_col_types = FALSE)
msv <- ms %>% dplyr::transmute(
  Cohort = subspecialty_abbrev,
  `Active stock` = format(n_active, big.mark = ","),
  `Departures/yr (primary)` = sprintf("%.1f", departures_primary),
  `Expected deaths/yr` = sprintf("%.1f", expected_mortality_per_yr),
  `Ratio (primary)` = sprintf("%.1f", ratio_primary),
  `Ratio (half missed)` = sprintf("%.1f", ratio_adj_half_missed),
  `Ratio (all missed)` = sprintf("%.1f", ratio_adj_all_missed),
  `2029 (primary)` = format(projected_2029_primary, big.mark = ","),
  `2029 (mortality-adj)` = format(projected_2029_mortality_adj, big.mark = ","))
kable(msv, caption = "**Appendix Table S25.** Age- and sex-specific expected-mortality sensitivity. Expected annual deaths in the active stock are computed from the SSA Period Life Table, 2020 (as used in the 2023 OASDI Trustees Report; q(x) by single year of age and sex) applied to each active physician's certification-anchored age and recorded sex, then added to the departure numerator. Because general-population mortality overstates mortality among practicing physicians (healthy-worker effect), the all-missed column is a conservative upper bound; the half-missed column reflects that some deaths are already captured administratively (NPPES deactivation, billing cessation). Both cohorts remain above one-for-one under every column.")
```

# Consistent-Definition Baseline Sensitivity (Reviewer #3)

```{r table-s26}
cb <- readr::read_csv(here::here("cliff","data","consistent_definition_baseline_sensitivity.csv"), show_col_types = FALSE)
cbv <- cb %>% dplyr::transmute(
  Cohort = subspecialty_abbrev,
  `Baseline (primary)` = format(baseline_primary, big.mark = ","),
  `Baseline (anchored)` = format(baseline_consistent, big.mark = ","),
  `Delta` = sprintf("%+d", baseline_delta),
  `Ratio (primary)` = sprintf("%.1f", ratio_primary),
  `Ratio (anchored)` = sprintf("%.1f", ratio_consistent),
  `2029 (primary)` = format(projected_2029_primary, big.mark = ","),
  `2029 (anchored)` = format(projected_2029_consistent, big.mark = ","))
kable(cbv, caption = "**Appendix Table S26.** Consistent-definition baseline sensitivity. The 2025 active stock is regenerated under the SAME non-Open-Payments anchored rule used for departure events (a physician is active if not classified as departed under the anchored rule by 2025), removing the stock-versus-flow definitional asymmetry. The anchored baseline is modestly larger (the stricter anchor classifies fewer physicians as departed), and the completion-to-departure ratios remain well above replacement, so pairing a broad cohorting baseline with an anchored departure definition does not materially bias the stock-flow balance.")
```

# Inactivity-Threshold Sensitivity (Reviewer #9)

```{r table-s27}
it <- readr::read_csv(here::here("cliff","data","inactivity_threshold_sensitivity.csv"), show_col_types = FALSE)
itv <- it %>% dplyr::transmute(
  Cohort = subspecialty_abbrev,
  `Inactivity threshold (yr)` = threshold_years,
  `Departure events (window)` = n_events_in_window,
  `Departure rate (%/yr)` = sprintf("%.2f", departure_rate_pct),
  `Mean annual departures` = sprintf("%.1f", avg_annual_departures),
  `Completion-to-departure ratio` = sprintf("%.2f", replacement_ratio),
  Assessment = assessment)
kable(itv, caption = "**Appendix Table S27.** Inactivity-duration sensitivity (distinct from the calendar-window sensitivity). A departure event is re-derived directly from Medicare Part B clinical activity: a physician whose last Part B active year is L is classified as departed in year L+1 only if the required T-year absence is fully observable by 2023 (L + T <= 2023); otherwise the exit is right-censored. Only the age-band hazard changes with T; the active baseline and entrant counts are held at the primary values. This isolates the Part B inactivity signal and is more aggressive than the multi-source consensus primary definition, so ratios are lower than the primary 5.6/7.1; nonetheless both board-certified cohorts remain above replacement at every threshold. The 2- and 3-year thresholds are indistinguishable within the fully-observable 2016-2021 window (both require the last active year to fall by 2020 for an in-window event), and a stricter 4-year threshold raises the ratio because fewer exits meet the longer absence requirement.")
```

# Hierarchical Partial-Pooling Hazard (Reviewer #4)

```{r table-s28}
hh <- readr::read_csv(here::here("cliff","data","hierarchical_hazard_comparison.csv"), show_col_types = FALSE)
mlab <- c(unpooled = "Unpooled (specialty-specific)", pooled = "Pooled (primary)",
          partial_pooled = "Partial-pooled (hierarchical)",
          `partial_pooled_abu_0.5x` = "Partial-pooled, ABU hazard x0.5",
          `partial_pooled_abu_2x` = "Partial-pooled, ABU hazard x2")
hhv <- hh %>% dplyr::transmute(
  Cohort = subspecialty_abbrev,
  Method = mlab[method],
  `Completion-to-departure ratio` = sprintf("%.2f", replacement_ratio),
  `95% interval` = ifelse(is.na(ci_lo), "-", sprintf("%.2f to %.2f", ci_lo, ci_hi)),
  Events = events, `Person-years` = person_years)
kable(hhv, caption = "**Appendix Table S28.** Hierarchical partial-pooling sensitivity for the age-band departure hazard. A binomial discrete-time hazard model with a shared age-band shape (fixed band effects) and a specialty deviation (random intercept) was fit by penalized maximum likelihood with a weakly-informative prior on the between-specialty variance (blme::bglmer); with only two board-certified cohorts, plain maximum likelihood collapses the specialty variance to zero (full pooling), so the variance prior is what makes this a genuine partial pool. Ratios are recomputed by applying each hazard set to the cohort's active-age distribution through the same dynamic model. Uncertainty: unpooled and pooled intervals are a 4,000-draw Beta(events+0.5, person-years-events+0.5) posterior on the band hazards; the partial-pooled interval is a 300-iteration parametric bootstrap of the fitted model (seed 20260720). The partial-pooled ratio lies between the unpooled and pooled estimates for each cohort, and all methods (with their intervals) remain above one-for-one replacement. ABU-hazard uncertainty is folded in by scaling the ABU share of the URPS hazard by 0.5 and 2.")
```

# Graduate-Supply Scenarios (Reviewer #8)

```{r table-s29}
gs <- readr::read_csv(here::here("cliff","data","graduate_growth_scenarios.csv"), show_col_types = FALSE)
gsv <- gs %>% dplyr::transmute(
  Cohort = subspecialty_abbrev, Scenario = scenario_label,
  `Annual graduates` = sprintf("%.0f", annual_graduates),
  `Completion-to-departure ratio` = sprintf("%.2f", replacement_ratio),
  `Projected 2029` = format(projected_2029, big.mark = ","), Assessment = assessment)
kable(gsv, caption = "**Appendix Table S29.** Graduate-supply scenarios beyond the flat four-year mean. The four observed ACGME completer counts (AY2020-21..2023-24) are recast into transparent forward assumptions and each is passed through the same dynamic model (only the annual entrant count changes; the pooled hazard and active baseline are held at primary values). Cohort accounting uses the most recent observed year (fellows already in final training, the most directly projectable near-term output); contraction uses the lowest of the four recent years; the cautious trend is the four-year mean plus half the observed linear annual trend over the horizon midpoint, capped at the highest observed year. The conservative (70% conversion) and optimistic (NRMP benchmark) scenarios are shown for reference. Both board-certified cohorts remain above replacement under every graduate-supply scenario.")
```

# Estimated-Baseline Lag Decomposition (Reviewer #5)

```{r table-s30}
bl <- readr::read_csv(here::here("cliff","data","baseline_lag_decomposition.csv"), show_col_types = FALSE)
blv <- bl %>% dplyr::transmute(
  Cohort = subspecialty_abbrev,
  `Directly supported (2023 Part B)` = directly_supported,
  `Carried forward 1-2 yr` = carried_fwd_1_2yr,
  `Carried forward >=3 yr` = carried_fwd_3plus,
  `No Part B history` = no_admin_history,
  `ABU net-new` = abu_net_new,
  `Estimated 2025 baseline` = format(baseline_total, big.mark = ","),
  `Ratio (primary)` = sprintf("%.2f", ratio_primary),
  `Ratio (long-lag removed)` = sprintf("%.2f", ratio_lag_trimmed))
kable(blv, caption = "**Appendix Table S30.** Decomposition of the estimated 2025 active baseline by administrative-support recency (Medicare Part B last-active year). The baseline is carried forward from admin data ending 2023-2024; the great majority of each Part B-observable cohort had Part B activity in the latest administrative year (2023), and only a small number rest on a three-year-or-longer carry-forward. The final column is a baseline-lag sensitivity: removing the long-lag (>=3-year carried-forward) physicians from the active stock, the most adverse plausible baseline-classification error, does not weaken the conclusion (the ratio is unchanged to slightly higher, because those physicians also carry projected departures). ABU net-new urogynecologists (URPS only) are recent certificants added to the both-pathway baseline; physicians with no Medicare Part B history (non-Medicare panels) are a residual limitation.")
```

# Age-Band-Stratified Temporal Back-Test (Reviewer #10)

```{r table-s31}
bt <- readr::read_csv(here::here("cliff","data","temporal_backtest_ageband.csv"), show_col_types = FALSE)
btv <- bt %>% dplyr::transmute(
  Cohort = subspecialty_abbrev, `Age band (2016)` = age_band,
  `N at base` = n_base, `Observed active 2021` = observed_active,
  `Predicted active 2021` = sprintf("%.1f", predicted_active),
  `Error (%)` = ifelse(is.na(pct_error), "-", sprintf("%+.1f", pct_error)))
kable(btv, caption = "**Appendix Table S31.** Age-band-stratified closed-cohort temporal back-test. The board-certified cohort active in 2016 is frozen and rolled forward to 2021 under the age-band hazard model (pure survival, no entrants); the predicted number still active is compared with the observed number within each base-year age band. Calibration is close in the well-populated bands (within a few percent through age 59, which hold the large majority of physicians); the oldest bands contain as few as four to nine physicians per cohort, so their percentage errors reflect small counts rather than model misfit. ABU net-new urogynecologists are absent from the historical back-test cohort (no pre-2024 roster) and are not included.")
```

# Data Availability

All input data sources are publicly available. Board certification, NPI, practice location, and Medicare and Open Payments records are identifiable professional records; reported analytic results are aggregated. The consolidated
workforce-projection table analyzed here is version-controlled with the analysis
code and is validated on load by a fail-closed data contract
(`manuscript/R/workforce_data_contract.R`); primary workforce-projection values are
generated programmatically from that consolidated dataset, and secondary sensitivity
analyses from their respective version-controlled analytic outputs.

# Supplemental References

<div id="refs"></div>
